---
title: "COVID-19 Policy Stringency and Outdoor Recreation: The Case of Resident Marine Sportfishing in the United States"
author:
  - name: David W. Carter
    email: david.w.carter@noaa.gov
    affiliation: NOAA Fisheries Southeast Fisheries Science Center
    footnote: Corresponding Author
  - name: Alexander Gordan
    email: alexander.gordan@noaa.gov
    affiliation: ECS
  - name: Sabrina Lovell
    email: sabrina.lovell@noaa.gov
    affiliation: NOAA Fisheries Office of Science and Technology
address:
  - code: NOAA Fisheries Southeast Fisheries Science Center
    address: 75 Virginia Beach Drive, Miami, FL, 33149
  - code: ECS
    address: Department 2, Street, City, State, Zip
  - code: NOAA Fisheries Office of Science and Technology
    address: Department 2, Street, City, State, Zip
header-includes:
   - \usepackage{threeparttable}    
abstract: |
  Governments responded to the Covid-19 pandemic with different policies to curtail the spread of the virus. We show how sportfishing levels are related to the stringency of Covid-19 policies. Specifically, the number of sportfishing trips bears an inverse-U-shaped relationship with Covid-19 policy stringency. 
  
author_summary: |
  He does this and that.
  
bibliography: C19PolicyRec.bib
output:
  bookdown::pdf_book:
    base_format: rticles::plos_article
csl: plos.csl
editor_options: 
  chunk_output_type: console
---

_Text based on plos sample manuscript, see <https://journals.plos.org/ploscompbiol/s/latex>_

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(feather)
library(survey)
options(survey.lonely.psu="adjust",multicore=TRUE)
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(texreg)
library(cdlTools)
library(kableExtra)
library(ciTools)
library(sandwich)
library(lmtest)
library(fst)
```

```{r,echo=FALSE,warning=FALSE,eval=FALSE}

#RUN ONCE TO CREATE DATASET: EVAL SET TO FALSE AFTER

#census_api_key("dc352be83efcd7e4189b8a5d1be652187e50c66d",install=TRUE)

#Load Atlantic and Gulf of Mexico MRIP data
trips=read_feather("trips1720.feather")
#Remove observations:
#without weights
#without months
#headboats
#non-residents 
#aggregate(I(ST_RES==ST)~MODE_FX+YEAR,FUN=mean,data=trips)
trips=subset(trips,wp_int>0 & month!=99 & MODE_FX!=4 & ST==ST_RES)

#Create counting variable for aggregation purposes
trips$one=1

#Mark party size of trips greather than 12 people as missing
#99% if of trips have less than 13 people
#cumsum(prop.table(table(trips$PARTY[trips$MODE_FX==7])))
#cumsum(prop.table(table(trips$PARTY[trips$MODE_FX==5])))
trips$PARTY[trips$PARTY>12]=NA

#Marking hours fished greater than 24 hours as missing
#table(trips$HRSF,useNA = "always")
#There were also a few trips with zero hours fished, marking as missing
trips$HRSF[trips$HRSF>24 | trips$HRSF==0]=NA

#Merge in distance dataset
dist=read_fst("distAll.fst")
trips=merge(trips,dist,by=c("ZIP","ST","CNTY","INTSITE"),all.x=TRUE)

tripsTM = 
  trips %>%
  complete(YEAR,month,MODE_FX,ST) %>%
  group_by(YEAR,month,MODE_FX,ST) %>%
  summarize(N_obs = n(),
            Trips = sum(wp_int,na.rm=TRUE),
            Hrsf  = weighted.mean(HRSF,wp_int,na.rm=TRUE),
            Party = weighted.mean(PARTY,wp_int,na.rm=TRUE),
            Miles = weighted.mean(MILES,wp_int,na.rm=TRUE),
            Htvl  = weighted.mean(TRAVELHOURS,wp_int,na.rm=TRUE)) 


#prop.table(table(TripsTM$N_obs==1 & !is.na(TripsTM$Hrsf)))


#Setting missing trip estimates to zero and preparing variables for merge
#tripsTM$trips[is.na(tripsTM$Trips)]=0
#tripsTM$HRSF[is.na(tripsTM$HRSF)]=0
#tripsTM$PARTY[is.na(tripsTM$PARTY)]=0


tripsTM$MONTH=as.numeric(tripsTM$month)
tripsTM$year=as.numeric(as.character(tripsTM$YEAR))
tripsTM$fips=as.numeric(as.character(tripsTM$ST))



#Function to retrieve annual state population estimates from ACS
acsY = function(y) {
  d=get_acs(geography = "state", 
            variables = "B01001_001", 
            year = y,
            survey = "acs1")
  d$year=y
  return(d)
}

#Retrieve state population estimates for 2017-2019
pop=do.call(rbind,lapply(2017:2019,FUN=function(x) acsY(x)))

#Get 2020 Census state population estimates and merge to the prior years
pop20 <- get_decennial(geography = "state", 
                       variables = "P1_001N", 
                       year = 2020)
pop20$year=2020
pop20=plyr::rename(pop20,replace=c("value" = "estimate"))
pop=plyr::rbind.fill(pop,pop20)
pop$fips=as.numeric(pop$GEOID)

#Merge the population estimates to the trip dataset year and state
tripsTMP=merge(tripsTM,pop,by.x=c("year","fips"),by.y=c("year","fips"),all.x=TRUE)

#Download daily stringency index for US states and reshape into a long dataset
#https://github.com/OxCGRT/USA-covid-policy
stringentw= openxlsx::read.xlsx(xlsxFile = "OxCGRTUS_timeseries_all.xlsx", sheet = 1)
stringent <- reshape::melt(stringentw, id=names(stringentw[1:5]))

#Create month, year, and state fips variables that are compatible withe the trip dataset
stringent$month=match(substr(stringent$variable,3,5),month.abb)
stringent$year=as.numeric(substr(stringent$variable,6,9))
stringent$fips=fips(stringent$region_name)

#Create an aggregated stringency dataset by summing the index by state, year, and month
stringentm=aggregate(value~year+month+fips,FUN=sum,data=stringent,na.rm=TRUE)

#Merge stringency index data with trip data by year, month, and state
tripsTMPS=merge(tripsTMP,stringentm,by.x=c("year","MONTH","fips"),by.y=c("year","month","fips"),all.x=TRUE)
tripsTMPS$value[is.na(tripsTMPS$value)]=0

#Cleaning up and naming variables in dataset
tripsTMPS=subset(tripsTMPS,select=-c(YEAR,MONTH,variable,moe,fips,ST))
names(tripsTMPS)=c("Year","Month","Mode","Obs","Trips","Hrsf","Party","Miles","Htvl",
                   "FIPS","State","Population","Stringency")

#Trips per capita
tripsTMPS$TPC=tripsTMPS$Trips/tripsTMPS$Population

#Save the final dataset
saveRDS(tripsTMPS,"tripsTMPS.rds")

```

```{r loadData,echo=FALSE,warning=FALSE}
#Load the final dataset
tripsTMPS=readRDS("tripsTMPS.rds")

```

# Introduction

Marine angling is a major outdoor activity in the US. According to the most recent Marine Angler Expenditure Report by @lovell2020economic, expenditures on marine angling trips totaled $10.5B, and the 2019 Outdoor Participation Report from the Outdoor Foundation @2019OutdoorReport finds that 4.3% of Americans participated in saltwater fishing in 2018. For  context, that is comparable to 4.9% of Americans engaged in skiing and 2.9% participated in mountain biking, while the most common outdoor activities are running, with 19.2% of Americans participating, and fishing (fresh or saltwater) at 16.4%.  Furthermore, aside from angling's importance as a source of outdoor recreation, some recreational fisheries are also substantial components of the broader fishery management system, particularly in the Southeastern US, where sportfishing can account for substantial portions of the harvest of certain commercially and culturally significant species such as reef fish species in the Gulf of Mexico.

The COVID-19 pandemic and ensuing public (and private) policy responses had far-reaching social and economic impacts, and marine angling is no exception. While other research has investigated the impacts of COVID on outdoor recreation @landry2021has and angling @midway2021covid more broadly, we are the first to look specifically at the impact of COVID and COVID policy on the overall level and kind of marine angling activity in the US.

At a broad level, there are several reasons to study the impact of COVID-19 on marine angling:

1. Overall impact on fishing effort: if COVID-19 has increased (decreased) fishing effort in a persistent fashion, particularly for areas and modes of fishing with many species managed by ACL's, then management may need to respond by tightening (lengthening) seasons.

2. New entrants: In addition to short-term changes in the level and kind of fishing activity, we are also interested in whether COVID may have induced new, and especially younger, participants to the fishery. This could potentially have more long term implications, if the longstanding trend of higher churn among younger fishing license holders^[This can be seen on a data dashboard maintained by the America Sportfishing Association, where for instance you can see that for Florida, the churn rate (% of license holders not renewing their prior year's license) for license holders below 35 years old has historically been notably higher than that for older participants, but this trend reversed in 2020, with the churn rate for the youngest participants decreasing from 67% to 37%, while the rate for the oldest participants increased slightly from around 50% to 53%: https://asafishing.org/data-dashboard/]  were to be overturned, and as a result more younger fishermen would become more avid and more skilled, potentially becoming long-term stakeholders in the fishery.

3. Changes in fishing motivation and satisfaction, which coincides with new entrants to the fishery and congestion effects.

4. Generalizable lessons: COVID-19 catalyzed many other socio-economic changes, in particular a shift toward remote work, increased internal migration and business formation, and possibly an increase in labor bargaining power. These trends seem likely to persist beyond the end of the pandemic, and will have implications for all aspects of public management, with fisheries management being no exception. COVID impacts thus need to be understood not just as limited to the pandemic era, but as harbingers of things to come. 

In this paper, we focus on the first item. In particular, we investigate the relationship between marine fishing trips and COVID 19 policy stringency. It has been largely understood that COVID 19 exerted a positive influence on outdoor recreation activity generally, and angling in particular. However, the pandemic ushered in a complex tangle of causal pathways, including some which dampened fishing activity, and catalyzed a broad range of social changes, many of which will persist well into the future. It is thus necessary to understand in greater detail the individual factors and causal pathways through which the pandemic bump in fishing activity operated if we wish to obtain general insights which can inform fisheries management going forward.


# Background

Landry et al (2021) provide a useful summary of potential causal pathways through which COVID-19, and the policy response to it, may affect outdoor recreation generally, which we adapt here for the specific instance of angling:

1. Site closures, which may force anglers to cancel trips if a boat ramp or pier was closed.^[A state-by-state summary of some of this information is available in this spreadsheet: https://docs.google.com/spreadsheets/d/12n_FR0Re2_tpL_oZCdDCqWSIZf5smGn5palKhhh-CYk]

2. Concern about infection risk, which may have led people to cancel planned trips, and cease to plan new trips.

3. An increase in leisure time, due to reduced commuting time and work-from-home, as well as increased unemployment^[though note that the latter involves an ambiguous income effect, depending on whether or not the person's pre-pandemic income was high enough that the loss of income was not offset by unemployment supplement under the CARES act.]

4.Decreased availability of substitute leisure activities, especially indoor activities, which may have made angling relatively more attractive as a way to spend one's leisure time.

To these considerations, we may also add:

5. Relocation, as it is known both from census data and cell phone location data that internal migration in the US picked up substantially during the pandemic, and it is likely that the Southeastern states in our sample saw an uptick in population, while the northern states (at least some of them) like Massachusetts may have seen a decrease in their population. Considering that anglers are people who are very sensitive to location, site quality, etc. this may be particularly salient for them, and the rate of internal migration for anglers may be even higher than for the general population (ooh, this is a good question for the NSAR data).

6. Increases in labor bargaining power, as labor shortages developed simultaneously with increased unemployment, there is evidence that employers have been moved to compete on non-pecuniary margins, such as offering employees greater flexibility in schedules, which may be advantageous to anglers in allowing them to make good on days with favorable fishing weather and get out when they may otherwise have had less flexibility to balance their work obligations with their hobby.





# Methods

## Data

To conduct our analysis, we construct a panel data set at the state-month level
spanning 2017 through 2020 for 16 US states, by merging two primary data
sources: (1) the Marine Recreational Information Program (MRIP), which provides
the necessary angler survey data to construct monthly estimates of the fishing activity in
each state, providing our dependent variable, and (2) the Oxford COVID
Government Response Tracker (OxCGRT), which provides fine-grained (daily)
variation in the level of COVID policy stringency in each of the US states,
which we collapse to the monthly level to provide our key independent variable.
Below, we describe in greater detail these two major data sources, as well as
some supplementary data used in the analysis,  and our methodology for
processing these data sources to produce our panel dataset.

### MRIP

The MRIP program oversees survey research to monitor the level of marine angling
activity in a consistent fashion throughout US fisheries (we discuss the structure of
these surveys in greater detail below). The primary unit of measurement for
our aggregated summary of the MRIP dataset is the angler-trip (hereafter
abbreviated simply to trip), representing the total number of distinct trips
multiplied by the average party size of those trips.

We use the MRIP data for the 16 states along the Gulf and Atlantic coasts in which
it is available. The MRIP program publishes estimates of
fishing activity for 2-month periods referred to as waves, but rather than use
these estimates we use the publicly-available microdata files^[The MRIP data
files are maintained by NOAA fisheries at the following URL:
https://www.fisheries.noaa.gov/recreational-fishing-data/recreational-fishing-data-downloads]
to estimate total trips in each state-month, and additionally we use other
information in those trip files to calculate the average hours fished per trip
and average party size on charter and private trip modes, which are
supplementary dependent variables for our analysis. Creating our own estimates
also allows us to customize the estimates, and we choose to specifically create
estimates of the number of trips taken by in-state residents.

We assemble monthly observations from 2017-2020 for three different modes of
fishing: private boats, charter boats, and shore. With `r
length(unique(tripsTMPS$Year))` years, `r length(unique(tripsTMPS$Month))`
months in each year and `r length(unique(tripsTMPS$State))` states, the data set
for each mode has `r length(tripsTMPS$Trips[tripsTMPS$Mode==7])` observations.

The MRIP program currently consists of 2 primary data collection efforts, the
Fishing Effort Survey (FES), and the Access Point Angler Intercept Survey
(APAIS). The FES is a mail survey which asks anglers to recall how many trips
they have been on recently, in order to get information on the total level of
fishing activity. The APAIS is an in-person survey conducted at common fishing
locations by field workers, and asks for information about the fishing trip they
are on when surveyed, as well as sampling some of the fish they catch in order
to get valid length information for the caught fish , as well as species
composition. During 2020, the year in which our treatment occurs, the FES was
able to continue unhindered, while the APAIS was suspended for at least some
length of time pretty much everywhere in March and April 2020, with data
collection resuming unevenly throughout the country in May and June (wave 3)
according to local conditions^[see Figure 6 of the MRIP survey design manual for
more information on which states resumed data collection at which times:
https://media.fisheries.noaa.gov/2021-09/MRIP-Survey-Design-and-Statistical-Methods-2021-09-15.pdf].
Therefore, while the total trip estimates are available for all state-months in
2020 in the same fashion as in 2019 and earlier years, the supplementary
information on average hours fished and party size has reduced availability, and
accordingly the sample size for these regressions are reduced due to this
missing data problem.


### COVID Stringency Index

Our key independent variable, on state-month level COVID stringency, comes from the Oxford COVID Government Response Tracker project. This data, as described by Hale et al (2021), is a global dataset tracking government response to COVID across 19 distinct policy indicators such as school- and work-closures, stay-at-home-orders, and movement restrictions. These data are tracked at the daily level, and for the US they are available by state. The data have been continuously updated since February 2020 by a team of volunteers who parse government reports, news, and other sources of info to create a standardized set of indicators which are comparable across the globe and across time. The full dataset includes indicators for economic response and health policy, as well as the closure indicators mentioned previously. The data product we use is the "Stringency Index", which combines all 8 closure indicators, as well as a variable indicating "Public information campaign", such as the "flatten the curve" campaign which was common throughout the US circa March and April 2020. All indicators in this index are recorded on discrete, ordinal scales with either 3, 4 or 5 ordinal levels. 
Hallas et al (2021) present additional information pertinent to the US state-level data, as well as the stylized facts in the data, such as the multiple cycles of first stringency tightening, followed by loosening policy until a new COVID wave and increased community transmission results in a re-tightening of COVID stringency. Additionally, over time the political gap in Stringency widened with democratic-led states having progressively tighter COVID policy compared to conservative states. 
We take the daily level data for each state and collapse it to a state-month level dataset by summing the value of the Stringency index for each day in the month. This data is then merged with the MRIP data for analysis.

### Supplementary Data

In order to make the MRIP data on a comparable basis across states, we calculate
per capita trip level using annual population estimates for each state bordering
the Atlantic and Gulf of Mexico. For 2017-2019, these are estimates from the
American Community Survey, and for 2020 we use figures from the Decennial
Census.

### Descriptive Statistics

```{r, echo=FALSE,warning=FALSE}
tpclt20=aggregate(TPC~State+Mode,FUN=mean,data=tripsTMPS,subset=Year<2020)
names(tpclt20)[3]="TPClt20"
#Average trips, stringency, and trips per capita by state and mode for 2020
mbms=aggregate(cbind(Trips,Stringency,TPC)~State+Mode,FUN=mean,data=tripsTMPS,subset=Year==2020)
#Combine the means for 2020 and years prior by state and mode
mbms=merge(mbms,tpclt20,by=c("State","Mode"))
#Ratio of the average trips per capita in 2020 to the average trips per capita in prior years by state and mode
mbms$TPCdTPClt20=mbms$TPC/mbms$TPClt20
mbms=mbms[order(mbms$Mode,mbms$State),]
# add state abbreviations
mbms <- mbms %>% 
  left_join(fips_codes %>% select(state, state_name), 
            by = c("State" = "state_name"))
```

```{r private-boat-trips-seasonality-plot,echo=FALSE,warning=FALSE,fig.height=6,fig.width=5,fig.align='center', fig.cap="2019 Monthly Marine Sportfishing Trips by State. The seasonal profile of each state is plotted as a separate grey line, with each month's value expressed as a percentage of the peak month. The bold line gives a simple average of these relative monthly values across the 16 states."}
tripsTMPS %>%
  filter(Mode == 7 & Year == 2019) %>%
  group_by(State) %>% 
  mutate(relative_TPC = TPC / max(TPC)) %>% 
  ggplot(aes(x=Month, y=relative_TPC, group=State)) +
  geom_line(alpha = 0.08) +
  stat_summary(mapping = aes(x=Month,  y=relative_TPC, group = 1), 
               fun = "mean", 
               geom = "line",
               inherit.aes = FALSE) +
  ylab("Trips (relative to peak trips)") +
  ggtitle("Private Boats")
```

Figure \@ref(fig:private-boat-trips-seasonality-plot) displays seasonal trends
in the level of trips across the 16 states in the dataset. Each grey line plots
the seasonal pattern for one state, with the values scaled relative to the state's
peak month. The black line provides a simple average of these relative values across
the 16 states within each month, so that the black line represents the composite
national-level seasonality trend in fishing trips among these 16 states. July is
both the modal peak month among the states, and the peak month for the national 
trend. There are many state-months in which 0 trips are recorded,^[The [MRIP Data 
User Handbook](https://media.fisheries.noaa.gov/2021-09/MRIP%20Data%20User%20Handbook%20Updated%202021-09-30.pdf) explains this in Table 1: many states are not sampled during January and February,
and Maine is not sampled in March or April either. In the [MRIP Survey Design Manual](https://media.fisheries.noaa.gov/2021-09/MRIP-Survey-Design-and-Statistical-Methods-2021-09-15.pdf), Figure 6 shows the state-month's which do not have coverage in white, along with detail of the 2020 data quality for state-months which are covered. So these should *not* be considered "true zeros", but rather non-randomly missing data, as the state-months which are not covered are such due to the presumption of low fishing activity attributable to unfavorable weather.] and in 
particular there are only 4 states with positive trips in January: Florida (with January trips equal to 43% of the peak value), North Carolina (37%), Alabama (27%), and
Mississippi (21%).

```{r charter-boat-trips-seasonality-plot,echo=FALSE,warning=FALSE,fig.height=6,fig.width=5,fig.align='center', fig.cap="2019 Monthly Marine Sportfishing Trips by State. The seasonal profile of each state is plotted as a separate grey line, with each month's value expressed as a percentage of the peak month. The bold line gives a simple average of these relative monthly values across the 16 states."}
tripsTMPS %>%
  filter(Mode == 5 & Year == 2019) %>%
  group_by(State) %>% 
  mutate(relative_TPC = TPC / max(TPC)) %>% 
  ggplot(aes(x=Month, y=relative_TPC, group=State)) +
  geom_line(alpha = 0.08) +
  stat_summary(mapping = aes(x=Month,  y=relative_TPC, group = 1), 
               fun = "mean", 
               geom = "line",
               inherit.aes = FALSE) +
  ylab("Trips (relative to peak trips)") +
  ggtitle("Charter Boats")
```

```{r shore-trips-seasonality-plot,echo=FALSE,warning=FALSE,fig.height=6,fig.width=5,fig.align='center', fig.cap="2019 Monthly Marine Sportfishing Trips by State. The seasonal profile of each state is plotted as a separate grey line, with each month's value expressed as a percentage of the peak month. The bold line gives a simple average of these relative monthly values across the 16 states."}
tripsTMPS %>%
  filter(Mode == 3 & Year == 2019) %>%
  group_by(State) %>% 
  mutate(relative_TPC = TPC / max(TPC)) %>% 
  ggplot(aes(x=Month, y=relative_TPC, group=State)) +
  geom_line(alpha = 0.08) +
  stat_summary(mapping = aes(x=Month,  y=relative_TPC, group = 1), 
               fun = "mean", 
               geom = "line",
               inherit.aes = FALSE) +
  ylab("Trips (relative to peak trips)") +
  ggtitle("Shore")
```

Figures \@ref(fig:charter-boat-trips-seasonality-plot) and
\@ref(fig:shore-trips-seasonality-plot) replicate Figure 
\@ref(fig:private-boat-trips-seasonality-plot), but for charter and shore mode
trips, respectively.


```{r private-boat-trips-seasonality-plot2,echo=FALSE,warning=FALSE,fig.height=6,fig.width=5,fig.align='center', fig.cap="2019 Monthly Marine Sportfishing Trips by State, relative to annual total."}
tripsTMPS %>%
  filter(Mode == 7 & Year == 2019) %>%
  group_by(State) %>% 
  mutate(relative_TPC = TPC / sum(TPC)) %>% 
  ggplot(aes(x=Month, y=relative_TPC, group=State)) +
  geom_line(alpha = 0.08) +
  stat_summary(mapping = aes(x=Month,  y=relative_TPC, group = 1), 
               fun = "mean", 
               geom = "line",
               inherit.aes = FALSE) +
  ylab("Trips (as a fraction of annual total)") +
  ggtitle("Private Boats")
```

Figure \@ref(fig:private-boat-trips-seasonality-plot2) shows the same curves as 
in Figure \@ref(fig:private-boat-trips-seasonality-plot), but rescaled to show
monthly values as a fraction of the annual total. This shows that fishing activity
is in most states fairly broadly distributed from April through November, with 
only 10 state-months accounting for more than 20% of the 
annual trips in a state, and a maximum of 35% for Maine in June. All but 1 of the 10
state-months with more than 20% of annual trips come from the Northern states of
Maine, New Hampshire, Massachusetts, Delaware and Rhode Island, with July in 
Mississippi as the exception.

```{r COVID-stringency-phillips-curve-FL,echo=FALSE,warning=FALSE,fig.cap="COVID Stringency vs relative fishing trips over time for Florida, private mode. Each label on the figure plots values for a single month, with sequential months connected by the path. The y-axis values are the trips for the month in 2020, divided by the average trips in that month from 2017 through 2019.",fig.height=6,fig.width=5,fig.align='center'}
tripsTMPS %>% 
  filter(Mode == 7) %>% 
  group_by(State, Month) %>% 
  arrange(Year, .by_group = TRUE) %>% 
  mutate(trip_ratio = case_when(Year == 2020 ~ Trips / mean(Trips[1:3]))) %>% 
  ungroup() %>% 
  arrange(Month) %>% 
  filter(State == "Florida", Year==2020) %>% 
  ggplot(aes(x = Stringency, y = trip_ratio, label = Month)) +
  ylab("Trip Ratio (2020:2017-19)") +
  geom_path() +
  geom_label()
```

Figure \@ref(fig:COVID-stringency-phillips-curve-FL) shows the comovement of
COVID Stringency and relative fishing activity for the state of Florida over the course of
2020. With the notable exception of February to March, most of the month-to-month
changes exhibit a downward slope. Note that Florida was chosen for this exhibit because it is the
only state with sufficient fishing activity throughout the entire year to create
these ratios for all 12 months of the year.

```{r pString,echo=FALSE,warning=FALSE,fig.cap="Monthly 2020 stringency Index by State",fig.height=6,fig.width=5,fig.align='center'}
tripsTMPS %>% 
  filter(Year > 2019) %>%
  ggplot( aes(x=Month,y=Stringency, group=State, color=State)) +
  geom_line() + 
  ylab("stringency")
```




```{r tripsVstring,echo=FALSE,warning=FALSE,fig.cap="Average Trips per Capita versus Covid-19 Policy Stringency",fig.height=5,fig.width=6,fig.align='center'}

with(subset(mbms,Mode==7),
     plot(TPCdTPClt20 ~Stringency,pch=NA,main="Private Boats",xlim=c(1000,2000),
          ylab="Trips per Capita Ratio",xlab="Covid-19 Policy Stringency"))
with(subset(mbms,Mode==7 ),
     text(TPCdTPClt20 ~Stringency,font=2,cex=0.75,
          labels=state[Mode==7 ]))

with(subset(mbms,Mode==5 ),
     plot(TPCdTPClt20 ~Stringency,pch=NA,main="Charter Boats",xlim=c(1000,2000),
          ylab="Trips per Capita Ratio",xlab="Covid-19 Policy Stringency"))
with(subset(mbms,Mode==5 ),
     text(TPCdTPClt20 ~Stringency,font=2,cex=0.75,
          labels=state[Mode==5 ]))

with(subset(mbms,Mode==3 ),
     plot(TPCdTPClt20 ~Stringency,pch=NA,main="Shore",xlim=c(1000,2000),
          ylab="Trips per Capita Ratio",xlab="Covid-19 Policy Stringency"))
with(subset(mbms,Mode==3 ),
     text(TPCdTPClt20 ~Stringency,font=2,cex=0.75,
          labels=state[Mode==3 ]))

```



## Modeling Approach

Quasi-poisson regression estimated with mean equal to $exp(pop)*exp(xb)$ and variance equal to $scale*mean$.



# Results

Our results represent a composite effect of the actual local COVID policies, as well as localized attitudes, risk tolerances, and beliefs about COVID-19 which are correlated with COVID stringency. In particular, we see the inverse-U relationship we find to be the outcome of 2 countervailing pathways (laid out in the background section), (1) stringency reduces available substitutes of indoor or high-density outdoor activities, increasing the attractiveness of fishing, which is the dominant force on the low-stringency side of the curve, and (2) COVID risk makes even a relatively distanced activity, such as fishing, less appealing compared to entirely in-home activities such as cooking, backyard games, video games, etc., which is the dominating force on the high-stringency side of the curve.

Overall, our moderate results for the relationship between COVID policy stringency and total angler-trips taken, combined with the null results for relationship with changes in tripp attributes like hours fished and party size, suggests that fishing activity (marine fishing activity) was actually rather stable throughout 2020, in comparison with other socioeconomic indicators like unemployment, GDP, etc. which saw 3-sigma or 4-sigma changes, sometimes the largest month-to-month changes since these statistics have been recorded.
This is to some degree consistent with another study, which conducted a nation-wide survey using the license frames from multiple states of both freshwater and saltwater anglers, and found that among the survey respondents, the modal response to the question "How has CVOID most impacted your life" was "No effect", followed by "stress". This suggests that anglers are perhaps less affected by COVID than the general population, which is known to have experienced record-levels of stress and anxiety during the COVID pandemic, according to research including a nationally-representative survey on mental health published in Science.
How to interpret this is an open question. It is commonly reported that angling is a relaxing activity, which many people do for solace from the world ("the contemplative man’s recreation"), and so perhaps it is the case that the kind of person who goes fishing is simply less affected by the pandemic. It may also reflect other correlations, such as occupational composition, if it is the case that anglers tend to have jobs which were less affected by the pandemic, or political correlations, if anglers happen to be more conservative / less inclined to believe that the pandemic was a cause to change their behavior. 




```{r sumSat,echo=FALSE,warning=FALSE,results='asis',message=FALSE}

# ssbym=as.data.frame(t(aggregate(cbind(Trips,Hrsf,Party,Stringency)~Mode,data=tripsTMPS,
#           FUN=function(x) cbind(mean(x), median(x), sd(x)))))
# 
# names(ssbym)=c("Shore","Charter","Private")
# rownames(ssbym)=c("Mode",rep(c("Mean","Median","St. Dev."),4))
# 
# kbl(ssbym[-1,],row.names = rep(c("Mean","Median","St. Dev."),4)) %>%
#   kable_classic() %>%
#   pack_rows("Trips", 1, 3) %>%
#   pack_rows("Hours Fished", 4, 6)  %>%
#   pack_rows("Party Size", 7, 9)  %>%
#   pack_rows("Stringency", 10, 12)

library(vtable)
st(tripsTMPS %>% 
     mutate(Mode = recode(Mode, `3` = "Shore", `5` = "Charter", `7` = "Private")) %>% 
     select("Trips","Hrsf","Party","Stringency", "Mode"),
   group = 'Mode', group.test = FALSE,group.long = TRUE,
   title="Summary Statistics for Key Variables, by Mode",
   out="latex")

```



```{r pfe,echo=FALSE,warning=FALSE}

#Fixed effect poisson regressions of trips on stringency index with a population offset
#Modeling trips per capita
mf=formula(Trips ~ State + Month + I(Stringency/100) + I((Stringency/100)^2) + offset(log(Population)))
pm=glm(mf,data=tripsTMPS,family = quasipoisson,subset=Mode==7 )
cm=glm(mf,data=tripsTMPS,family = quasipoisson,subset=Mode==5 )
sm=glm(mf,data=tripsTMPS,family = quasipoisson,subset=Mode==3 )

```

```{r regTbl,echo=FALSE,warning=FALSE,results='asis'}

staten=gsub("State","",names(coef(pm)))[2:16]

texreg(list(pm,cm,sm),custom.model.names = c("Private","Charter","Shore"),single.row = FALSE,digits=3,label="regResults",
       caption = "Quasi-Poisson Fixed Effect Regression of Trips on Covid-19 Stringency, by Mode",
       #custom.coef.names = c("Intercept",staten,month.name[-1],
       #                       "Stringency/100","(Stringency/100)^2"),
       custom.coef.map=list("(Intercept)" = "Intercept",
                            "I(Stringency/100)"="Stringency/100","I((Stringency/100)^2)"="(Stringency/100)^2"),
       include.aic=FALSE,include.bic=FALSE,include.loglik=FALSE,caption.above = TRUE,
       custom.note = "%stars. The base state is Alabama and the base Month is January.",
       custom.gof.rows = list("Scale" = c(summary(pm)$dispersion,summary(cm)$dispersion,summary(sm)$dispersion)))


```


```{r tripStringRel,echo=FALSE,warning=FALSE,fig.cap="Predicted Trips as a Function of the Stringency Index",fig.height=8,fig.width=4}

#Dataset to predict trips as a function of stringency
#holding everything else constant.
#Using June in Florida: zero stringency should correspond with
#the mean historical trips in Florida during June
nd=data.frame(State="Florida",Month="06",
              Population=with(tripsTMPS,mean(Population[State=="Florida" & Month=="06"])),
              Stringency=seq(0,3000,by=50))

#Predicted trips as a function of stringency, including confidence intervals
pd=add_ci(nd,pm)
cd=add_ci(nd,cm)
sd=add_ci(nd,sm)

#Plot predicted trips and confidence intervals
par(mfrow=c(3,1))

plot(y=pd$pred/1000000,x=pd$Stringency,type="l",ylim=c(2,5),
     main="Private Boats",
     xlab="Stringency Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Millions of Fishing Trips (Scaled to June in FL)")
lines(y=pd$LCB0.025/1000000,x=pd$Stringency,lty=2)
lines(y=pd$UCB0.975/1000000,x=pd$Stringency,lty=2)

plot(y=cd$pred/1000000,x=cd$Stringency,type="l",ylim=c(.04,.09),
     main="Charter Boats",
     xlab="Stringency Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Millions of Fishing Trips (Scaled to June in FL)")
lines(y=cd$LCB0.025/1000000,x=cd$Stringency,lty=2)
lines(y=cd$UCB0.975/1000000,x=cd$Stringency,lty=2)

plot(y=sd$pred/1000000,x=sd$Stringency,type="l",ylim=c(3,7),
     main="Shore",
     xlab="Stringency Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Millions of Fishing Trips (Scaled to June in FL)")
lines(y=sd$LCB0.025/1000000,x=sd$Stringency,lty=2)
lines(y=sd$UCB0.975/1000000,x=sd$Stringency,lty=2)

par(mfrow=c(1,1))

#Use this to reverse the x-axis: xlim=c(3000,0)

#Don't think that we need robust standard error with the
#Quasi-poisson, but here is how you would calculate them
#coeftest(pm, vcov = vcovHC(pm, type="HC4"))

#Bootstrapped prediction CI to verify (takes long to run as currently coded)
# library(boot)
# 
# bf=function(d,i){
#   pm=glm(mf,data=d[i,],family = quasipoisson )
#   pd=predict(pm,newdata = nd,type="response")/1000000
#   return(pd)
# }
# 
# bcp=boot(subset(tripsTMPS,Mode==7),bf,R=1000)
# zz=sapply(1:61,FUN=function(i) boot.ci(boot.out = bcp, type = c("bca"),index=i)$bca)
# 
# plot(pd,type="l",ylim=c(2,6))
# lines(zz[4,],lty=2)
# lines(zz[5,],lty=2)

```


## Discussion

Having written all this, let us now turn again to the eternal question: what are the policy implications??
If internal migration has increased, perhaps certain states in the Southeast need to take account of increased fishing pressure, while other states can ease off of their attempts to constrain effort. If increases in congestion have effectively increased the cost of trips, then perhaps angler welfare per trip is reduced, which on a naive interpretation would mean that their allocation of quota should be reduced, although the other way to look at that is that it points to an increased need for flexibility in disposition of catch privileges. In other words, if boat ramp congestion due to increased boating activity (non-fishing related) is negatively affecting anglers and eating into the value of trip, anglers may be better off being able to choose their own seasons to adapt to their local conditions, which they will be able to do better than a central planner. 


# References {#references .unnumbered}

<div id="refs"></div>







