---
title: "COVID-19 Policy Stringency and Outdoor Recreation: The Case of Resident Marine Sportfishing in the United States"
author:
  - name: David W. Carter
    email: david.w.carter@noaa.gov
    affiliation: NOAA Fisheries Southeast Fisheries Science Center
    footnote: Corresponding Author
  - name: Alexander Gordan
    email: alexander.gordan@noaa.gov
    affiliation: ECS
  - name: Sabrina Lovell
    email: sabrina.lovell@noaa.gov
    affiliation: NOAA Fisheries Office of Science and Technology
address:
  - code: NOAA Fisheries Southeast Fisheries Science Center
    address: 75 Virginia Beach Drive, Miami, FL, 33149
  - code: ECS
    address: Department 2, Street, City, State, Zip
  - code: NOAA Fisheries Office of Science and Technology
    address: Department 2, Street, City, State, Zip
header-includes:
   - \usepackage{threeparttable}    
abstract: |
  Governments responded to the Covid-19 pandemic with different policies to curtail the spread of the virus. We show how sportfishing levels are related to the stringency of Covid-19 policies. Specifically, the number of sportfishing trips bears an inverse-U-shaped relationship with Covid-19 policy stringency. 
  
author_summary: |
  He does this and that.
  
bibliography: C19PolicyRec.bib
output:
  bookdown::pdf_book:
    base_format: rticles::plos_article
csl: plos.csl
editor_options: 
  chunk_output_type: console
---

_Text based on plos sample manuscript, see <https://journals.plos.org/ploscompbiol/s/latex>_

```{r setup0,echo=FALSE}

knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)

```


```{r setup}

library(feather)
library(survey)
options(survey.lonely.psu="adjust",multicore=TRUE)
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(texreg)
library(cdlTools)
library(kableExtra)
library(ciTools)
library(sandwich)
library(lmtest)
library(fst)

library(fixest)
library(car)
library(clubSandwich)
library(timetk)
library(lubridate)

library(ggpubr)
library(ggeffects)

```

```{r,eval=FALSE}

#RUN ONCE TO CREATE DATASET: EVAL SET TO FALSE AFTER

trips <- read_feather("data/trips1721.feather")

tripsTM <- trips %>% 
  filter(wp_int > 0, month != 99, MODE_FX != 4, ST == ST_RES) %>% 
  # exclude anomalous observations
  mutate(PARTY = if_else(PARTY > 12, NA_real_, PARTY),
         HRSF = if_else((HRSF > 24) | (HRSF == 0), NA_real_, HRSF)) %>% 
  complete(YEAR,month,MODE_FX,ST) %>%
  group_by(YEAR,month,MODE_FX,ST) %>%
  summarize(N_obs = n(),
            Trips = sum(wp_int,na.rm=TRUE),
            Hrsf  = weighted.mean(HRSF,wp_int,na.rm=TRUE),
            Party = weighted.mean(PARTY,wp_int,na.rm=TRUE),
            .groups = "drop") %>% 
  mutate(MONTH = as.numeric(month),
         year = as.numeric(as.character(YEAR)),
         fips = as.numeric(as.character(ST)))

#Function to retrieve annual state population estimates from ACS
acsY = function(y) {
  d=get_acs(geography = "state", 
            variables = "B01001_001", 
            year = y,
            survey = "acs1")
  d$year=y
  return(d)
}

#Retrieve state population estimates for 2017-2019
pop=do.call(rbind,lapply(2017:2019,FUN=function(x) acsY(x)))
pop

pop20 <- get_decennial(geography = "state", 
                       variables = "P1_001N", 
                       year = 2020)

NST_EST2021_alldata <- read_csv("data/NST-EST2021-alldata.csv")

pop21 <- NST_EST2021_alldata %>% 
  rename(fips = STATE, Population = POPESTIMATE2021) %>% 
  filter(fips != "00") %>% 
  select(fips, Population) %>% 
  mutate(year = 2021)


pop20 <- pop20 %>% 
  rename(fips = GEOID, Population = value) %>% 
  select(fips, Population) %>% 
  mutate(year = 2020)

pop <- pop %>% 
  rename(fips = GEOID, Population = estimate) %>% 
  select(fips, Population, year) %>% 
  bind_rows(pop20) %>% 
  bind_rows(pop21) %>% 
  mutate(fips = as.integer(fips))

tripsTMP <- tripsTM %>% 
  left_join(pop, by = c("fips", "year"))

fips_codes_state <- fips_codes %>% select(state_code, state_name) %>% distinct()

stringency <- read_csv("data/OxCGRT_US_latest.csv") %>% 
  rename(State = RegionName) %>% 
  mutate(Year = as.integer((substr(Date, start = 1, stop = 4))),
         Month = substr(Date, start = 5, stop = 6),
         Day = substr(Date, start = 7, stop = 8)) %>% 
  # next 2 commands assign the last known value of Stringency for days with NA
  arrange(State, Year, Month, Day) %>% 
  fill(StringencyIndex) %>% 
  group_by(State, Year, Month) %>%
  summarise(Stringency = mean(StringencyIndex),
            .groups = "drop") %>% 
  left_join(fips_codes_state, by = c("State" = "state_name")) %>% 
  rename(fips = state_code) %>% 
  mutate(fips = as.integer(fips))

fips_codes_state_alt <- fips_codes_state %>% mutate(state_code = as.numeric(state_code))

tripsTMPS <- tripsTMP %>% 
  left_join(stringency, by = c("YEAR" = "Year", "month" = "Month", "fips" = "fips")) %>% 
  rename(Year = YEAR,
         Month = month,
         Mode= MODE_FX,
         Obs = N_obs,
         FIPS = fips) %>% 
  select(!State) %>% 
  left_join(fips_codes_state_alt, by = c("FIPS" = "state_code")) %>% 
  rename(State = state_name) %>% 
  select(Year,Month,Mode,Obs,Trips,Hrsf,Party,FIPS,State,Population,Stringency) %>% 
  mutate(TPC = Trips / Population,
         Stringency = if_else(is.na(Stringency), 0, Stringency),
         TripsR=round(Trips,0),
         StateF=as.factor(gsub(" ","",State)),
         StateF=relevel(StateF,ref="Florida"),
         MonthF=as.factor(as.numeric(Month)),
         MonthF=relevel(MonthF,ref="6"),
         YearF=as.factor(as.numeric(Year)),
         YearF=relevel(YearF,ref="2019"),
         sm=interaction(StateF,MonthF),
         sm=relevel(sm,ref="Florida.6"),
         ym=interaction(as.factor(Year),MonthF),
         ym=relevel(ym,ref="2019.6"),
         Year19s=Year-2019,
         date = make_date(Year, Month,1))

saveRDS(tripsTMPS, "tripsTMPS.rds")

```

```{r loadData,echo=FALSE,warning=FALSE}
#Load the final dataset
tripsTMPS=readRDS("tripsTMPS.rds")

no_data_sms<- tripsTMPS %>% 
  filter(Mode == "7") %>% 
  group_by(State, Month) %>% 
  summarise(max_Trips = max(Trips), .groups = "drop") %>% 
  filter(max_Trips == 0) %>% 
  select(State, Month)

data_for_new_model <- tripsTMPS %>% 
  anti_join(no_data_sms, by = c("State", "Month"))

no_data_sms5<- tripsTMPS %>%
  filter(Mode == "5") %>%
  group_by(State, Month) %>%
  summarise(max_Trips = max(Trips), .groups = "drop") %>%
  filter(max_Trips == 0) %>%
  select(State, Month)

data_for_new_model5 <- tripsTMPS %>%
  anti_join(no_data_sms5, by = c("State", "Month"))

```

# Introduction

Saltwater angling is a major outdoor activity in the U.S. with expenditures on fishing trips on totaling more than $10 billion @lovell2020economic. Over 4% of Americans participate in saltwater fishing compared with around around 20% for running, the most common outdoor recreation activity @2019OutdoorReport.  Furthermore, in some marine fisheries, recreational angling is a substantial component of harvest, especially in the Southeastern US, where sportfishing can account for a large share commercially and culturally significant species @lewin2019potential.

The COVID-19 pandemic and ensuing policy responses had far-reaching social and economic implications, and marine angling is no exception. Landry et. al. @landry2021has summarize the potential pathways through which COVID-19 and related policy responses may affect outdoor recreation. For recreational fishing the potential pathways are as follows:

1. Site closures may force anglers to cancel trips if, for example, boat ramps or piers were closed.

2. Concern about infection risk may have led people to cancel planned trips or cease to plan new trips.

3. People may have more leisure time due to work-from-home flexibility, reduced commuting time, or job loss.

4. A decrease in substitute leisure activities, especially indoor activities may have made angling relatively more attractive as a way to spend one's leisure time [@midway2021covid; @morales2021contrasting]

The first two pathways imply a decrease in fishing effort while the latter two imply an increase. @link2021noaa documents interruptions in U.S. marine fishery data collections due to COVID-19 policies such as lockdowns. Many of these interruptions also affected anglers access to fishing launch sites.^[Our focus is on saltwater anglers, but @paradis2021can found that around ninety percent of freshwater fishing jurisdictions in North America were able to keep fishing open during spring of 2020]. In a national survey of U.S. anglers, most respondents experienced some access restrictions, but nearly all respondents reported that they did not think that recreational fishing was unsafe during the early stages of the pandemic in 2020 @midway2021covid. The same study found that anglers reported very little change in the number of trips they took relative to the same period in years past. However, this is self-reported changes for one period in time averaged over various levels of COVID-19 policy stringencies.  Actual changes in effort over a longer period of time and range of COVID-19 policy stringencies may be different.     

The net effect of the four pathways depends on prevailing conditions at the time effort is measured. For example, early in the response to the pandemic when lockdowns where common, the effect of site closures and concerns about infection risk would dominate and effort should be relatively low. At other times, however, as lockdowns are relaxed, but restrictions on indoor gatherings and leisure travel remain, effort could increase due the effects of increased leisure time and lack of substitutes for leisure time. A similar type of variation in trip taking behavior could happen over space, i.e. the different levels of COVID-19 policy stringency in different jurisdictions (e.g., states) could contribute to variations in angler trip taking behavior across jurisdictions.

We use data on the estimated monthly number of saltwater fishing trips from states on the U.S. east coast along with a monthly, state-level measure of COVID-19 policy stringency to examine the relationship between the stringency of COVID-19 policy and the  level of marine angling activity. Our results suggest that the relation ship is nonlinear whereby the number of fishing trips increases at a decreasing rate as COVID-19 policy becomes more stringent. This finding is important to understanding how people respond to measures aimed at containing virus spread and can help in planning during future pandemics.


<!-- To these considerations, we may also add: -->

<!-- 5. Relocation, as it is known both from census data and cell phone location data that internal migration in the US picked up substantially during the pandemic, and it is likely that the Southeastern states in our sample saw an uptick in population, while the northern states (at least some of them) like Massachusetts may have seen a decrease in their population. Considering that anglers are people who are very sensitive to location, site quality, etc. this may be particularly salient for them, and the rate of internal migration for anglers may be even higher than for the general population (ooh, this is a good question for the NSAR data). -->

<!-- 6. Increases in labor bargaining power, as labor shortages developed simultaneously with increased unemployment, there is evidence that employers have been moved to compete on non-pecuniary margins, such as offering employees greater flexibility in schedules, which may be advantageous to anglers in allowing them to make good on days with favorable fishing weather and get out when they may otherwise have had less flexibility to balance their work obligations with their hobby. -->

In this paper, we measure the net effect of stringency policies on recreational effort over the range of stringency implemented accross the states in the eastern U.S. In particular, we investigate the relationship between marine fishing trips and COVID 19 policy stringency. It has been largely understood that COVID 19 exerted a positive influence on outdoor recreation activity generally, and angling in particular. However, the pandemic ushered in a complex tangle of causal pathways, including some which dampened fishing activity, and catalyzed a broad range of social changes, many of which will persist well into the future. It is thus necessary to understand in greater detail the individual factors and causal pathways through which the pandemic bump in fishing activity operated if we wish to obtain general insights which can inform fisheries management going forward. Specifically, if COVID-19 has changed fishing effort in a persistent fashion then management may need to respond with different regulations and management strategies in the future.

<!-- There are several reasons to study the impact of COVID-19 on marine angling: -->

<!-- 1. Overall impact on fishing effort: if COVID-19 has increased (decreased) fishing effort in a persistent fashion, particularly for areas and modes of fishing with many species managed by ACL's, then management may need to respond by tightening (lengthening) seasons. -->

<!-- 2. New entrants: In addition to short-term changes in the level and kind of fishing activity, we are also interested in whether COVID may have induced new, and especially younger, participants to the fishery. This could potentially have more long term implications, if the longstanding trend of higher churn among younger fishing license holders^[This can be seen on a data dashboard maintained by the America Sportfishing Association, where for instance you can see that for Florida, the churn rate (% of license holders not renewing their prior year's license) for license holders below 35 years old has historically been notably higher than that for older participants, but this trend reversed in 2020, with the churn rate for the youngest participants decreasing from 67% to 37%, while the rate for the oldest participants increased slightly from around 50% to 53%: https://asafishing.org/data-dashboard/]  were to be overturned, and as a result more younger fishermen would become more avid and more skilled, potentially becoming long-term stakeholders in the fishery. -->

<!-- 3. Changes in fishing motivation and satisfaction, which coincides with new entrants to the fishery and congestion effects. -->

<!-- 4. Generalizable lessons: COVID-19 catalyzed many other socio-economic changes, in particular a shift toward remote work, increased internal migration and business formation, and possibly an increase in labor bargaining power. These trends seem likely to persist beyond the end of the pandemic, and will have implications for all aspects of public management, with fisheries management being no exception. COVID impacts thus need to be understood not just as limited to the pandemic era, but as harbingers of things to come. -->

# Methods

## Data

To conduct our analysis, we construct a panel data set at the state-month level spanning 2017 through 2020 for 16 US states, by merging two primary data sources: (1) the Marine Recreational Information Program (MRIP), which provides the necessary angler survey data to construct monthly estimates of the fishing activity in each state, providing our dependent variable, and (2) the Oxford COVID Government Response Tracker (OxCGRT), which provides fine-grained (daily) variation in the level of COVID policy stringency in each of the US states, which we collapse to the monthly level to provide our key independent variable.
Below, we describe in greater detail these two major data sources, as well as some supplementary data used in the analysis,  and our methodology for processing these data sources to produce our panel dataset.

### MRIP

The MRIP program oversees survey research to monitor the level of marine angling activity in a consistent fashion throughout US fisheries (we discuss the structure of these surveys in greater detail below). The primary unit of measurement for our aggregated summary of the MRIP dataset is the angler-trip (hereafter abbreviated simply to trip), representing the total number of distinct trips multiplied by the average party size of those trips.

We use the MRIP data for the 16 states along the Gulf and Atlantic coasts in which it is available. The MRIP program publishes estimates of fishing activity for 2-month periods referred to as waves, but rather than use these estimates we use the publicly-available microdata files^[The MRIP data files are maintained by NOAA fisheries at the following URL:
https://www.fisheries.noaa.gov/recreational-fishing-data/recreational-fishing-data-downloads] to estimate total trips in each state-month. Creating our own estimates also allows us to customize the estimates, and we choose to specifically create
estimates of the number of trips taken by in-state residents.

We assemble monthly observations from 2017-2020 for three different modes of fishing: private boats, charter boats, and shore. With `r length(unique(tripsTMPS$Year))` years, `r length(unique(tripsTMPS$Month))` months in each year and `r length(unique(tripsTMPS$State))` states, the data set for each mode has `r length(tripsTMPS$Trips[tripsTMPS$Mode==7])` observations.

The MRIP program currently consists of 2 primary data collection efforts, theFishing Effort Survey (FES), and the Access Point Angler Intercept Survey (APAIS). The FES is a mail survey which asks anglers to recall how many trips they have been on recently, in order to get information on the total level of fishing activity. The APAIS is an in-person survey conducted at common fishing
locations by field workers, and asks for information about the fishing trip they are on when surveyed, as well as sampling some of the fish they catch in order to get valid length information for the caught fish , as well as species composition. During 2020, the year in which our treatment occurs, the FES was able to continue unhindered, while the APAIS was suspended for at least some
length of time nationwide in March and April 2020, with data collection resuming unevenly throughout the country in May and June (wave 3) according to local conditions^[see Figure 6 of the MRIP survey design manual for more information on which states resumed data collection at which times: https://media.fisheries.noaa.gov/2021-09/MRIP-Survey-Design-and-Statistical-Methods-2021-09-15.pdf.
We are able to construct the total trip estimates are available for all state-months in 2020 in the same fashion as in 2019 and earlier years, since the FES was able to continue unhindered. However, other outcome variables which might have been investigated,
such as the average party size on a fishing trip, are affected by the APAIS suspension and can not be estimated for many
state-months in early 2020, which is when the largest changes in COVID-19 policy stringency occur, and therefore we can not examine the relationship between policy stringency and these outcomes due to intractable missing data problems.


### COVID Stringency Index

Our key independent variable, on state-month level COVID stringency, comes from the Oxford COVID Government Response Tracker project. This data, as described by Hale et al (2021), is a global dataset tracking government response to COVID across 19 distinct policy indicators such as school- and work-closures, stay-at-home-orders, and movement restrictions. These data are tracked at the daily level, and for the US they are available by state. The data have been continuously updated since February 2020 by a team of volunteers who parse government reports, news, and other sources of info to create a standardized set of indicators which are comparable across the globe and across time. The full dataset includes indicators for economic response and health policy, as well as the closure indicators mentioned previously. The data product we use is the "Stringency Index", which combines all 8 closure indicators, as well as a variable indicating "Public information campaign", such as the "flatten the curve" campaign which was common throughout the US circa March and April 2020. All indicators in this index are recorded on discrete, ordinal scales with either 3, 4 or 5 ordinal levels. 
Hallas et al (2021) present additional information pertinent to the US state-level data, as well as the stylized facts in the data, such as the multiple cycles of first stringency tightening, followed by loosening policy until a new COVID wave and increased community transmission results in a re-tightening of COVID stringency. Additionally, over time the political gap in Stringency widened with democratic-led states having progressively tighter COVID policy compared to conservative states. 
We take the daily level data for each state and collapse it to a state-month level dataset by summing the value of the Stringency index for each day in the month. This data is then merged with the MRIP data for analysis.

### Supplementary Data

In order to make the MRIP data on a comparable basis across states, we calculate per capita trip level using annual population estimates for each state bordering the Atlantic and Gulf of Mexico. For 2017-2019, these are estimates from the American Community Survey, and for 2020 we use figures from the Decennial Census.

### Descriptive Statistics

```{r, echo=FALSE,warning=FALSE}
tpclt20=aggregate(TPC~State+Mode,FUN=mean,data=tripsTMPS,subset=Year<2020)
names(tpclt20)[3]="TPClt20"
#Average trips, stringency, and trips per capita by state and mode for 2020
mbms=aggregate(cbind(Trips,Stringency,TPC)~State+Mode,FUN=mean,data=tripsTMPS,subset=Year==2020)
#Combine the means for 2020 and years prior by state and mode
mbms=merge(mbms,tpclt20,by=c("State","Mode"))
#Ratio of the average trips per capita in 2020 to the average trips per capita in prior years by state and mode
mbms$TPCdTPClt20=mbms$TPC/mbms$TPClt20
mbms=mbms[order(mbms$Mode,mbms$State),]
# add state abbreviations
mbms <- mbms %>% 
  left_join(fips_codes %>% select(state, state_name), 
            by = c("State" = "state_name"))
```

```{r private-boat-trips-seasonality-plot,fig.height=6,fig.width=5,fig.align='center', fig.cap="2019 Monthly Marine Sportfishing Trips by State. The seasonal profile of each state is plotted as a separate grey line, with each month's value expressed as a percentage of the peak month. The bold line gives a simple average of these relative monthly values across the 16 states."}
tripsTMPS %>%
  filter(Mode == 7 & Year == 2019) %>%
  group_by(State) %>% 
  mutate(relative_TPC = TPC / max(TPC)) %>% 
  ggplot(aes(x=Month, y=relative_TPC, group=State)) +
  geom_line(alpha = 0.08) +
  stat_summary(mapping = aes(x=Month,  y=relative_TPC, group = 1), 
               fun = "mean", 
               geom = "line",
               inherit.aes = FALSE) +
  ylab("Trips (relative to peak trips)") +
  ggtitle("Private Boats")
```

Figure \@ref(fig:private-boat-trips-seasonality-plot) displays seasonal trends
in the level of trips across the 16 states in the dataset. Each grey line plots
the seasonal pattern for one state, with the values scaled relative to the state's
peak month. The black line provides a simple average of these relative values across
the 16 states within each month, so that the black line represents the composite
national-level seasonality trend in fishing trips among these 16 states. July is
both the modal peak month among the states, and the peak month for the national 
trend. There are many state-months in which 0 trips are recorded,^[The [MRIP Data 
User Handbook](https://media.fisheries.noaa.gov/2021-09/MRIP%20Data%20User%20Handbook%20Updated%202021-09-30.pdf) explains this in Table 1: many states are not sampled during January and February,
and Maine is not sampled in March or April either. In the [MRIP Survey Design Manual](https://media.fisheries.noaa.gov/2021-09/MRIP-Survey-Design-and-Statistical-Methods-2021-09-15.pdf), Figure 6 shows the state-month's which do not have coverage in white, along with detail of the 2020 data quality for state-months which are covered. So these should *not* be considered "true zeros", but rather non-randomly missing data, as the state-months which are not covered are such due to the presumption of low fishing activity attributable to unfavorable weather.] and in 
particular there are only 4 states with positive trips in January: Florida (with January trips equal to 43% of the peak value), North Carolina (37%), Alabama (27%), and
Mississippi (21%).

```{r charter-boat-trips-seasonality-plot,fig.height=6,fig.width=5,fig.align='center', fig.cap="2019 Monthly Marine Sportfishing Trips by State. The seasonal profile of each state is plotted as a separate grey line, with each month's value expressed as a percentage of the peak month. The bold line gives a simple average of these relative monthly values across the 16 states."}
tripsTMPS %>%
  filter(Mode == 5 & Year == 2019) %>%
  group_by(State) %>% 
  mutate(relative_TPC = TPC / max(TPC)) %>% 
  ggplot(aes(x=Month, y=relative_TPC, group=State)) +
  geom_line(alpha = 0.08) +
  stat_summary(mapping = aes(x=Month,  y=relative_TPC, group = 1), 
               fun = "mean", 
               geom = "line",
               inherit.aes = FALSE) +
  ylab("Trips (relative to peak trips)") +
  ggtitle("Charter Boats")
```

```{r shore-trips-seasonality-plot,fig.height=6,fig.width=5,fig.align='center', fig.cap="2019 Monthly Marine Sportfishing Trips by State. The seasonal profile of each state is plotted as a separate grey line, with each month's value expressed as a percentage of the peak month. The bold line gives a simple average of these relative monthly values across the 16 states."}
tripsTMPS %>%
  filter(Mode == 3 & Year == 2019) %>%
  group_by(State) %>% 
  mutate(relative_TPC = TPC / max(TPC)) %>% 
  ggplot(aes(x=Month, y=relative_TPC, group=State)) +
  geom_line(alpha = 0.08) +
  stat_summary(mapping = aes(x=Month,  y=relative_TPC, group = 1), 
               fun = "mean", 
               geom = "line",
               inherit.aes = FALSE) +
  ylab("Trips (relative to peak trips)") +
  ggtitle("Shore")
```

Figures \@ref(fig:charter-boat-trips-seasonality-plot) and
\@ref(fig:shore-trips-seasonality-plot) replicate Figure 
\@ref(fig:private-boat-trips-seasonality-plot), but for charter and shore mode
trips, respectively.


```{r trips-seasonality-plot,fig.height=6,fig.width=5,fig.align='center', fig.cap="2019 Monthly Marine Sportfishing Trips by State, relative to annual total. The seasonal profile of each state is plotted as a separate grey line, with each month's value expressed as a percentage of the total annual trips for the state. The bold line gives a simple average of these relative monthly values across the 16 states."}
mode_names <- c(
                    `3` = "Shore Trips",
                    `5` = "Charter Trips",
                    `7` = "Private Trips"
                    )

tripsTMPS %>%
  filter(Year == 2019) %>%
  group_by(State, Mode) %>% 
  mutate(relative_TPC = TPC / sum(TPC)) %>% 
  ggplot(aes(x=Month, y=relative_TPC, group=State)) +
  geom_line(alpha = 0.08) +
  stat_summary(mapping = aes(x=Month,  y=relative_TPC, group = 1), 
               fun = "mean", 
               geom = "line",
               inherit.aes = FALSE) +
  ylab("Trips (as a fraction of annual total)") +
  facet_wrap(vars(Mode), scales = "free", nrow = 3, labeller = as_labeller(mode_names))
```

Figure \@ref(fig:trips-seasonality-plot) shows the same curves as 
in Figure \@ref(fig:private-boat-trips-seasonality-plot), but rescaled to show
monthly values as a fraction of the annual total. This shows that fishing activity
is in most states fairly broadly distributed from April through November, with 
only 10 state-months accounting for more than 20% of the 
annual trips in a state, and a maximum of 35% for Maine in June. All but 1 of the 10
state-months with more than 20% of annual trips come from the Northern states of
Maine, New Hampshire, Massachusetts, Delaware and Rhode Island, with July in 
Mississippi as the exception.

```{r COVID-stringency-phillips-curve-FL,fig.cap="COVID Stringency vs relative fishing trips over time for Florida, private mode. Each label on the figure plots values for a single month, with sequential months connected by the path. The y-axis values are the trips for the month in 2020, divided by the average trips in that month from 2017 through 2019.",fig.height=6,fig.width=5,fig.align='center'}
tripsTMPS %>% 
  filter(Mode == 7) %>% 
  group_by(State, Month) %>% 
  arrange(Year, .by_group = TRUE) %>% 
  mutate(trip_ratio = case_when(Year == 2020 ~ Trips / mean(Trips[1:3]))) %>% 
  ungroup() %>% 
  arrange(Month) %>% 
  filter(State == "Florida", Year==2020) %>% 
  ggplot(aes(x = Stringency, y = trip_ratio, label = Month)) +
  ylab("Trip Ratio (2020:2017-19)") +
  geom_path() +
  geom_label()
```

Figure \@ref(fig:COVID-stringency-phillips-curve-FL) shows the comovement of
COVID Stringency and relative fishing activity for the state of Florida over the course of
2020. With the notable exception of February to March, most of the month-to-month
changes exhibit a downward slope. Note that Florida was chosen for this exhibit because it is the
only state with sufficient fishing activity throughout the entire year to create
these ratios for all 12 months of the year.

```{r pString,fig.cap="Monthly 2020 stringency Index by State",fig.height=6,fig.width=5,fig.align='center'}
covid_colors <- c("NY" = "red", "FL" = "blue", "AL" = "darkgreen", "other" = "grey")
covid_alpha <- c("NY" = 1, "FL" = 1, "AL" = 1, "other" = 0.5)
tripsTMPS %>% 
  mutate(covid_cat = case_when(State == "New York" ~ "NY",
                               State == "Florida" ~ "FL",
                               State == "Alabama" ~ "AL",
                               TRUE ~ "other")) %>% 
  filter(Year > 2019) %>%
  mutate(Date = lubridate::make_date(Year, Month)) %>% 
  ggplot(aes(x=Date,y=Stringency, group=State, color=covid_cat, alpha=covid_cat)) +
  geom_line() + 
  ylab("stringency") + 
  scale_color_manual(values = covid_colors, name = "States") +
  scale_alpha_manual(values = covid_alpha, guide = "none")
```

Figure \@ref(fig:pString) shows the evolution of COVID policy stringency over the
course of 2020 for each of the 16 states in our analysis. Following the initial 
uniform jumps in stringency during the March/April 2020 lockdowns, there is 
substantial variation between the states in how COVID policy evolved. Here we have
highlighted 3 states showing arch-typical cases: (1) New York, which had 
consistently stringent COVID policy, (2) Alabama, where after April COVID policy was
loosened substantially, and stayed loose, and (3) Florida, which displays the most
variation in COVID policy over time. The substantial independent variation in these
series affords the opportunity to examine the relationship between COVID policy
stringency and fishing activity.

```{r decompPs, fig.cap="Seasonality by State: Private",fig.height=6,fig.width=5,fig.align='center'}


dSM=data_for_new_model %>%
  group_by(Mode,State) %>%
  tk_stl_diagnostics(date, box_cox_vec(TripsR),.message = FALSE)

pDecompS=function(m,t) {
  
  zz=subset(dSM,Mode==m)
  
  par(mfrow=c(4,4))
  
  if(t=="season") {
    sapply(unique(zz$State),FUN=function(x) plot(zz$season[zz$State==x],main=x,type="l"))
    } else {
      sapply(unique(zz$State),FUN=function(x) plot(zz$trend[zz$State==x],main=x,type="l"))
      }

par(mfrow=c(1,1))

}

pDecompS(7,"season")

```


```{r decompPt, fig.cap="Trend by State: Private",fig.height=6,fig.width=5,fig.align='center'}
pDecompS(7,"trend")
```

```{r decompCs, fig.cap="Seasonality by State: Charter",fig.height=6,fig.width=5,fig.align='center'}
pDecompS(5,"season")
```

```{r decompCt, fig.cap="Trend by State: Charter",fig.height=6,fig.width=5,fig.align='center'}
pDecompS(5,"trend")
```

```{r decompSs, fig.cap="Seasonality by State: Shore",fig.height=6,fig.width=5,fig.align='center'}
pDecompS(3,"season")
```

```{r decompSt, fig.cap="Trend by State: Shore",fig.height=6,fig.width=5,fig.align='center'}
pDecompS(3,"trend")
```

## Modeling Approach

We model the relationship between the aggregate monthly number of salwater fishing trips trips taken in each state and the monthly COVID policy stringency as a Poisson regression model with the following conditional mean:

$\mu_{im} = W_{it}*\exp(\alpha + \beta S_{it} + \lambda S_{it}^2 + \sum_{g=1}^{G} \sum_{m=1}^{M} z_{itg} d_{itm} \gamma_{itgm} + \sum_{y=1}^{Y} v_{ity}\delta_{ity})$

where for state $i$ during period $t$, $\mu_{it}$ is the expected aggregate trips, $W_{it}$ is the population, $S_{it}$ is the daily average COVID policy stringency. The dummy variables $d_{im}$ equals one for if the observation is from state  

<!-- Hellerstein, D.M., 1991. Using count data models in travel cost analysis with aggregate data. American journal of agricultural economics, 73(3), pp.860-866. -->

This model accounts for the count nature of the outcome variable, number
of trips taken, like a conventional Poisson model, but unlike a Poisson model it
allows for over-dispersion of the data @VerHoef2007. Specifically, the conditional mean is 
modeled as $\mu_i = \exp(pop_i)*\exp(x_i \beta)$, where $x_i$ is a vector of
explanatory variables for state-month $i$, $\beta$ is a vector of slope coefficients
to be estimated, and $pop_i$ is the population of state-month $i$; the variance
is modeled as $\mu_i * \theta$, where $\theta > 1$ is a scalar parameter to be
estimated which allows for the over-dispersion in the data and is also referred
to a scale parameter. 

In our specification, the vector $x_i$ includes 4 components: a linear term
for the COVID stringency index, a quadratic term for the COVID stringency index,
a set of state fixed effects to account for average differences in per capita
fishing trips between the states, and a set of month fixed effects to account for
the nation-wide average seasonality trend in fishing activity. 



# Results

Table \@ref(regResults) and Figure \@ref(fig:tripStringRel) present the results of
our estimated Quasi-Poisson regression model. A common trend is seen between
each of the 3 fitted models for the different fishing modes: There is an 
inverse-U relationship between COVID policy stringency and fishing trips, ie.
the first order relationship is an increase in fishing trips at moderate levels 
of policy stringency, with a second order effect dampening or even reversing
this relationship at higher levels of stringency. These relationships are seen
in the model coefficients presented in Table \@ref(regResults), which are significant
at the 5% level for private and shore modes, but not for charter. 

Figure \@ref(fig:tripStringRel)
represents these same relationships graphically, by plotting the estimated
model's predictions for the number of resident fishing trips that would occur in Florida
in the month of June with a given level of COVID stringency. The solid line
provides the point estimates for predicted trips, while the dotted lines plot
the 95% confidence intervals around these predictions. Note that while we 
present this relationship with the outcome scaled to the level of June in
Florida, the shape of the relationship between stringency and fishing trips
would be the same for any state-month combination.

Our results represent a composite effect of the actual local COVID policies, as
well as localized attitudes, risk tolerances, and beliefs about COVID-19 which
are correlated with COVID stringency. In particular, we see the inverse-U
relationship we find to be the outcome of 2 countervailing pathways (laid out in
the background section), (1) stringency reduces available substitutes of indoor
or high-density outdoor activities, increasing the attractiveness of fishing,
which is the dominant force on the low-stringency side of the curve, and (2)
COVID risk makes even a relatively distanced activity, such as fishing, less
appealing compared to entirely in-home activities such as cooking, backyard
games, video games, etc., which is the dominating force on the high-stringency
side of the curve.






```{r sumSat,results='asis',eval=FALSE}

# ssbym=as.data.frame(t(aggregate(cbind(Trips,Hrsf,Party,Stringency)~Mode,data=tripsTMPS,
#           FUN=function(x) cbind(mean(x), median(x), sd(x)))))
# 
# names(ssbym)=c("Shore","Charter","Private")
# rownames(ssbym)=c("Mode",rep(c("Mean","Median","St. Dev."),4))
# 
# kbl(ssbym[-1,],row.names = rep(c("Mean","Median","St. Dev."),4)) %>%
#   kable_classic() %>%
#   pack_rows("Trips", 1, 3) %>%
#   pack_rows("Hours Fished", 4, 6)  %>%
#   pack_rows("Party Size", 7, 9)  %>%
#   pack_rows("Stringency", 10, 12)

library(vtable)
st(tripsTMPS %>% 
     mutate(Mode = recode(Mode, `3` = "Shore", `5` = "Charter", `7` = "Private")) %>% 
     select("Trips","Hrsf","Party","Stringency", "Mode"),
   group = 'Mode', group.test = FALSE,group.long = TRUE,
   title="Summary Statistics for Key Variables, by Mode",
   out="latex")

```




```{r pfe,eval=FALSE}

rm=function(m,mn,y){
  
  if(m==5) {
    bd=data_for_new_model5
  } else {
      bd=data_for_new_model
    }

  d=subset(bd,Mode==m & Year %in% y)
  
  #Fixed effect poisson with:
  #Cluster robust standard errors with small-sample (small number of clusters) adjustment
  #Pustejovsky, J.E. and Tipton, E., 2018. Small-sample methods for cluster-robust
  #variance estimation and hypothesis testing in fixed effects models. Journal of Business
  #& Economic Statistics, 36(4), pp.672-683.
  
  #Baseline two-way fixed effect specification
  r0=fepois(TripsR ~ Stringency + I(Stringency^2) | StateF + YearF^MonthF ,
           data=d,offset = ~log(Population),
           vcov = function(x) vcovCR(x, type = "CR2",test = "Satterthwaite",cluster=d$StateF))   
  
  r= fepois(TripsR ~ Stringency + I(Stringency^2) | StateF^MonthF + StateF[Year19s] ,
           data=d,offset = ~log(Population),
           vcov = function(x) vcovCR(x, type = "CR2",test = "Satterthwaite",cluster=d$StateF)) 
  
  
  #Dataset to predict trips as a function of stringency
  #holding everything else constant.
  #Using June 2019 in Florida: zero stringency should correspond with
  #the predicted trips for this period in absence of stringency policies
  #Function to calculate the predicted trips
  #and associated confidence intervals using the 
  #delta method
  #The base prediction is for Wave 6 in Florida during 2019
  pDM=function(m,pop,x) {
    fe=fixef(m)
    feB=fe$`StateF^MonthF`["Florida_6"] + fe$StateF["Florida"]
    f=paste0(pop,"*exp(",feB," + ",x,"*Stringency + `I(Stringency^2)`*",x,"^2)")
    deltaMethod(m$coefficients,f,vcov=m$cov.scaled)}
  
  #Population in Wave 6 in Florida during 2019
  popE=d$Population[d$StateF=="Florida" & d$Year19s==0 & d$MonthF==6]

  p=t(sapply(0:100,function(x) pDM(r,popE,x)))
  p=as.data.frame(apply(p,2,unlist))
  
  pp=ggplot(p, aes(x = 0:100, y = Estimate, group = 1)) + 
    geom_line(col='black') + 
    geom_ribbon(aes(ymin = `2.5 %`, ymax = `97.5 %`), alpha = 0.1) +
    labs(title = mn,
         y="Expected June 2019 Florida Trips",
         x="Stringency") +
    geom_hline(yintercept=p[1,"Estimate"]) +
    scale_y_continuous(labels = scales::label_number(suffix = " M", scale = 1e-6))
  
  #Scaling the trip predictions to be relative to the base period
  ps=p
  ps[,"2.5 %"]=100*(ps[,"2.5 %"]-ps[1,"Estimate"])/ps[1,"Estimate"]
  ps[,"97.5 %"]=100*(ps[,"97.5 %"]-ps[1,"Estimate"])/ps[1,"Estimate"]
  ps[,"Estimate"]=100*(ps[,"Estimate"]-ps[1,"Estimate"])/ps[1,"Estimate"]
  
  pps=ggplot(ps, aes(x = 0:100, y = Estimate, group = 1)) + 
    geom_line(col='black') + 
    geom_ribbon(aes(ymin = `2.5 %`, ymax = `97.5 %`), alpha = 0.1) +
    labs(title = mn,
         y="% Difference from Expected June 2019 Florida Trips ",
         x="Stringency") +
    geom_hline(yintercept=0)
  
  #Elasticity calculations
  pme=t(sapply(0:100,FUN=function(x) 
    deltaMethod(r$coefficients,paste0(x,"*(Stringency + `I(Stringency^2)`*2*",x,")"),vcov=r$cov.scaled)))
  pme=as.data.frame(apply(pme,2,unlist))
  
  ppe=ggplot(pme, aes(x = 0:100, y = Estimate, group = 1)) + 
    geom_line(col='black') + 
    geom_ribbon(aes(ymin = `2.5 %`, ymax = `97.5 %`), alpha = 0.1) +
    labs(title = mn,
         y="Trip Elasticity",
         x="Stringency") +
    geom_hline(yintercept=0)
  
  
  return(list(reg0=r0,reg=r,pred=p,preds=ps,pplot=pp,pplots=pps,elast=pme,elastP=ppe))
  
}

pm=rm(7,"Private Boats",2017:2021)
cm=rm(5,"Charter Boats",2017:2021)
sm=rm(3,"Shore",2017:2021)

#Saving model and prediction objects because the predictions take long to run.
saveRDS(pm,"pm.rds")
saveRDS(cm,"cm.rds")
saveRDS(sm,"sm.rds")

```

```{r regTbl,results='asis'}

#Reading model and prediction objects
pm=readRDS("pm.rds")
cm=readRDS("cm.rds")
sm=readRDS("sm.rds")

texreg(list(pm$reg,cm$reg,sm$reg),custom.model.names = c("Private","Charter","Shore"),
       label="regResultsSS",caption.above = TRUE,digits=4,include.deviance=FALSE,include.bic=TRUE,
        caption = "Poisson Fixed Effect Regression of Trips on Covid-19 Stringency, by Mode",include.groups=FALSE,
       custom.note = "%stars. SEs clustered by state with small-sample adjustment.")

texreg(list(pm$reg0,cm$reg0,sm$reg0),custom.model.names = c("Private","Charter","Shore"),include.groups=FALSE,
       label="regResultsSS0",caption.above = TRUE,digits=4,include.deviance=FALSE,include.bic=TRUE,
       caption = "Poisson Fixed Effect Regression of Trips on Covid-19 Stringency, by Mode (2-way FE)",
       custom.note = "%stars. SEs clustered by state with small-sample adjustment.")

```


```{r tripStringRelSC,fig.cap="Predicted Trips as a Function of the Stringency Index: Levels",fig.height=8,fig.width=4}

#Predicted trips as a function of stringency, including confidence intervals
pa=ggarrange(pm$pplot,cm$pplot,sm$pplot,ncol = 1, nrow = 3)
annotate_figure(pa,fig.lab = "Note: Based on Trips in Florida During June of 2019",fig.lab.pos = "bottom.left")

```

```{r tripStringRelSS,fig.cap="Predicted Trips as a Function of the Stringency Index",fig.height=8,fig.width=4}

#Predicted trips as a function of stringency, including confidence intervals
pas=ggarrange(pm$pplots,cm$pplots,sm$pplots,ncol = 1, nrow = 3)
annotate_figure(pas,fig.lab = "Note: Based on Trips in Florida During June of 2019",fig.lab.pos = "bottom.left")

```

```{r tripStringElast,fig.cap="Elasticity as a Function of the Stringency Index",fig.height=8,fig.width=4}

#Elasticity as a function of stringency, including confidence intervals
pe=ggarrange(pm$elastP,cm$elastP,sm$elastP,ncol = 1, nrow = 3)

```

## Discussion

Overall, our moderate results for the relationship between COVID policy stringency and total angler-trips taken suggest that marine fishing activity was actually relatively stable throughout 2020, in comparison with other socioeconomic indicators like unemployment, GDP, etc. which saw 3- or 4-sigma changes, sometimes the largest month-to-month changes since these statistics have been recorded.
This is to some degree consistent with another study @midway2021covid, which conducted a nation-wide survey using the license frames from multiple states of both freshwater and saltwater anglers, and found that among the survey respondents, the modal response to the question "How has CVOID most impacted your life" was "No effect", followed by "stress" (Figure 2, panel b in the paper). This suggests that anglers are perhaps less affected by COVID than the general population, which is known to have experienced record-levels of stress and anxiety during the COVID pandemic, according to research including a nationally-representative survey on mental health published in Science @holman2020unfolding.
How to interpret this is an open question. It is commonly reported that angling is a relaxing activity, which many people do for solace from the world ("the contemplative man’s recreation"), and so perhaps it is the case that the kind of person who goes fishing is simply less affected by the pandemic. It may also reflect other correlations, such as occupational composition, if it is the case that anglers tend to have jobs which were less affected by the pandemic, or political correlations, if anglers happen to be more conservative / less inclined to believe that the pandemic was a cause to change their behavior. 

In regards to policy implications, our results may be interpreted in a few ways.
If internal migration has increased, perhaps certain states in the Southeast need to take account of increased fishing pressure, while other states can ease off of their attempts to constrain effort. If increases in congestion have effectively increased the cost of trips, then perhaps angler welfare per trip is reduced, which on a naive interpretation would mean that their allocation of quota should be reduced, although the other way to look at that is that it points to an increased need for flexibility in disposition of catch privileges. In other words, if boat ramp congestion due to increased boating activity (non-fishing related) is negatively affecting anglers and eating into the value of trip, anglers may be better off being able to choose their own seasons to adapt to their local conditions, which they will be able to do better than a central planner. 


# References {#references .unnumbered}

<div id="refs"></div>







