---
title: "COVID-19 Policy Stringency and Outdoor Recreation: The Case of Resident Marine Sportfishing in the United States"
author:
  - name: David W. Carter
    email: david.w.carter@noaa.gov
    affiliation: NOAA Fisheries Southeast Fisheries Science Center
    footnote: Corresponding Author
  - name: Alexander Gordan
    email: alexander.gordan@noaa.gov
    affiliation: ECS
  - name: Sabrina Lovell
    email: sabrina.lovell@noaa.gov
    affiliation: NOAA Fisheries Office of Science and Technology
address:
  - code: NOAA Fisheries Southeast Fisheries Science Center
    address: 75 Virginia Beach Drive, Miami, FL, 33149
  - code: ECS
    address: Department 2, Street, City, State, Zip
  - code: NOAA Fisheries Office of Science and Technology
    address: Department 2, Street, City, State, Zip
header-includes:
   - \usepackage{threeparttable}    
abstract: |
  Governments responded to the Covid-19 pandemic with different policies to curtail the spread of the virus. We show how sportfishing levels are related to the stringency of Covid-19 policies. Specifically, the number of sportfishing trips bears an inverse-U-shaped relationship with Covid-19 policy stringency. 
  
author_summary: |
  He does this and that.
  
bibliography: C19PolicyRec.bib
output:
  bookdown::pdf_book:
    base_format: rticles::plos_article
csl: plos.csl
editor_options: 
  chunk_output_type: console
---

_Text based on plos sample manuscript, see <https://journals.plos.org/ploscompbiol/s/latex>_

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(feather)
library(survey)
options(survey.lonely.psu="adjust",multicore=TRUE)
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(texreg)
library(cdlTools)
library(kableExtra)
library(ciTools)
library(sandwich)
library(lmtest)
library(fst)
```

```{r,echo=FALSE,warning=FALSE,eval=FALSE}

#RUN ONCE TO CREATE DATASET: EVAL SET TO FALSE AFTER

#census_api_key("dc352be83efcd7e4189b8a5d1be652187e50c66d",install=TRUE)

#Load Atlantic and Gulf of Mexico MRIP data
trips=read_feather("trips1720.feather")
#Remove observations:
#without weights
#without months
#headboats
#non-residents 
#aggregate(I(ST_RES==ST)~MODE_FX+YEAR,FUN=mean,data=trips)
trips=subset(trips,wp_int>0 & month!=99 & MODE_FX!=4 & ST==ST_RES)

#Create counting variable for aggregation purposes
trips$one=1

#Mark party size of trips greather than 12 people as missing
#99% if of trips have less than 13 people
#cumsum(prop.table(table(trips$PARTY[trips$MODE_FX==7])))
#cumsum(prop.table(table(trips$PARTY[trips$MODE_FX==5])))
trips$PARTY[trips$PARTY>12]=NA

#Marking hours fished greater than 24 hours as missing
#table(trips$HRSF,useNA = "always")
#There were also a few trips with zero hours fished, marking as missing
trips$HRSF[trips$HRSF>24 | trips$HRSF==0]=NA

#Merge in distance dataset
dist=read_fst("distAll.fst")
trips=merge(trips,dist,by=c("ZIP","ST","CNTY","INTSITE"),all.x=TRUE)

tripsTM = 
  trips %>%
  complete(YEAR,month,MODE_FX,ST) %>%
  group_by(YEAR,month,MODE_FX,ST) %>%
  summarize(N_obs = n(),
            Trips = sum(wp_int,na.rm=TRUE),
            Hrsf  = weighted.mean(HRSF,wp_int,na.rm=TRUE),
            Party = weighted.mean(PARTY,wp_int,na.rm=TRUE),
            Miles = weighted.mean(MILES,wp_int,na.rm=TRUE),
            Htvl  = weighted.mean(TRAVELHOURS,wp_int,na.rm=TRUE)) 


#prop.table(table(TripsTM$N_obs==1 & !is.na(TripsTM$Hrsf)))


#Setting missing trip estimates to zero and preparing variables for merge
#tripsTM$trips[is.na(tripsTM$Trips)]=0
#tripsTM$HRSF[is.na(tripsTM$HRSF)]=0
#tripsTM$PARTY[is.na(tripsTM$PARTY)]=0


tripsTM$MONTH=as.numeric(tripsTM$month)
tripsTM$year=as.numeric(as.character(tripsTM$YEAR))
tripsTM$fips=as.numeric(as.character(tripsTM$ST))



#Function to retrieve annual state population estimates from ACS
acsY = function(y) {
  d=get_acs(geography = "state", 
            variables = "B01001_001", 
            year = y,
            survey = "acs1")
  d$year=y
  return(d)
}

#Retrieve state population estimates for 2017-2019
pop=do.call(rbind,lapply(2017:2019,FUN=function(x) acsY(x)))

#Get 2020 Census state population estimates and merge to the prior years
pop20 <- get_decennial(geography = "state", 
                       variables = "P1_001N", 
                       year = 2020)
pop20$year=2020
pop20=plyr::rename(pop20,replace=c("value" = "estimate"))
pop=plyr::rbind.fill(pop,pop20)
pop$fips=as.numeric(pop$GEOID)

#Merge the population estimates to the trip dataset year and state
tripsTMP=merge(tripsTM,pop,by.x=c("year","fips"),by.y=c("year","fips"),all.x=TRUE)

#Download daily stringency index for US states and reshape into a long dataset
#https://github.com/OxCGRT/USA-covid-policy
stringentw= openxlsx::read.xlsx(xlsxFile = "OxCGRTUS_timeseries_all.xlsx", sheet = 1)
stringent <- reshape::melt(stringentw, id=names(stringentw[1:5]))

#Create month, year, and state fips variables that are compatible withe the trip dataset
stringent$month=match(substr(stringent$variable,3,5),month.abb)
stringent$year=as.numeric(substr(stringent$variable,6,9))
stringent$fips=fips(stringent$region_name)

#Create an aggregated stringency dataset by summing the index by state, year, and month
stringentm=aggregate(value~year+month+fips,FUN=sum,data=stringent,na.rm=TRUE)

#Merge stringency index data with trip data by year, month, and state
tripsTMPS=merge(tripsTMP,stringentm,by.x=c("year","MONTH","fips"),by.y=c("year","month","fips"),all.x=TRUE)
tripsTMPS$value[is.na(tripsTMPS$value)]=0

#Cleaning up and naming variables in dataset
tripsTMPS=subset(tripsTMPS,select=-c(YEAR,MONTH,variable,moe,fips,ST))
names(tripsTMPS)=c("Year","Month","Mode","Obs","Trips","Hrsf","Party","Miles","Htvl",
                   "FIPS","State","Population","Stringency")

#Trips per capita
tripsTMPS$TPC=tripsTMPS$Trips/tripsTMPS$Population

#Save the final dataset
saveRDS(tripsTMPS,"tripsTMPS.rds")

```

```{r loadData,echo=FALSE,warning=FALSE}
#Load the final dataset
tripsTMPS=readRDS("tripsTMPS.rds")

```

# Introduction

* Intro to C-19 and the anecdotal evidence related to C-19 and recreation
* Literature review
* Research question: How do outdoor recreation levels change as governments institute policies to protect public health during a pandemic?
* Approach: variation in state stringency policies to describe the relationship between C-19 and sportfishing levels

Landry et al (2021) provide a useful summary of potential causal pathways through which COVID-19 may affect outdoor recreation:

1. blah

2. Blah blah

3. Blah blah blah

To which we may add:

4. Etc

5. Etc etc




# Methods

## Data

We assemble monthly observations from 2017-2020 for three different modes of fishing: private boats, charter boats, and shore. With `r length(unique(tripsTMPS$Year))` years, `r length(unique(tripsTMPS$Month))` months in each year and `r length(unique(tripsTMPS$State))` states, the data set for each mode has `r length(tripsTMPS$Trips[tripsTMPS$Mode==7])` observations. 

* MRIP monthly estimates (via program) by mode (private, charter, shore) for each state bordering the Atlantic and Gulf of Mexico
  + refer to MRIP website with template programs
  + potential issue with PSEs
  + Only residents (show percentage of res. vs nonres.)

* Annual population estimates for each state bordering the Atlantic and Gulf of Mexico
  + 2017-2019: ACS
  + 2020 Census
  
* Monthly C-19 Stringency index for each state bordering the Atlantic and Gulf of Mexico
  + daily estimates summed to each month
  + higher numbers equate with more stringent policies


```{r, echo=FALSE,warning=FALSE}
tpclt20=aggregate(TPC~State+Mode,FUN=mean,data=tripsTMPS,subset=Year<2020)
names(tpclt20)[3]="TPClt20"
#Average trips, stringency, and trips per capita by state and mode for 2020
mbms=aggregate(cbind(Trips,Stringency,TPC)~State+Mode,FUN=mean,data=tripsTMPS,subset=Year==2020)
#Combine the means for 2020 and years prior by state and mode
mbms=merge(mbms,tpclt20,by=c("State","Mode"))
#Ratio of the average trips per capita in 2020 to the average trips per capita in prior years by state and mode
mbms$TPCdTPClt20=mbms$TPC/mbms$TPClt20
mbms=mbms[order(mbms$Mode,mbms$State),]
```

```{r private-boat-trips-seasonality-plot,echo=FALSE,warning=FALSE,fig.height=6,fig.width=5,fig.align='center', fig.cap="2019 Monthly Marine Sportfishing Trips by State. The seasonal profile of each state is plotted as a separate grey line, with each month's value expressed as a percentage of the peak month. The bold line gives a simple average of these relative monthly values across the 16 states."}
tripsTMPS %>%
  filter(Mode == 7 & Year == 2019) %>%
  group_by(State) %>% 
  mutate(relative_TPC = TPC / max(TPC)) %>% 
  ggplot(aes(x=Month, y=relative_TPC, group=State)) +
  geom_line(alpha = 0.08) +
  stat_summary(mapping = aes(x=Month,  y=relative_TPC, group = 1), 
               fun = "mean", 
               geom = "line",
               inherit.aes = FALSE) +
  ylab("Trips (relative to peak trips)") +
  ggtitle("Private Boats")
```

Figure \@ref(fig:private-boat-trips-seasonality-plot) displays seasonal trends
in the level of trips across the 16 states in the dataset. Each grey line plots
the seasonal pattern for one state, with the values scaled relative to the state's
peak month. The black line provides a simple average of these relative values across
the 16 states within each month, so that the black line represents the composite
national-level seasonality trend in fishing trips among these 16 states. July is
both the modal peak month among the states, and the peak month for the national 
trend. There are many state-months in which 0 trips occur,^[or at least no trips
are recorded by MRIP, need to point to the data documentation here] and in 
particular there are only 4 states with positive trips in January: Florida (with
 of trips occurring in January), North Carolina (37%), Alabama (27%), and
Mississippi (21%).^[why does South Carolina have 0 trips if North Carolina has 
positive trips and the ostensible explanation for the 0 trips is cold??]

```{r private-boat-trips-seasonality-plot2,echo=FALSE,warning=FALSE,fig.height=6,fig.width=5,fig.align='center', fig.cap="2019 Monthly Marine Sportfishing Trips by State, relative to annual total."}
tripsTMPS %>%
  filter(Mode == 7 & Year == 2019) %>%
  group_by(State) %>% 
  mutate(relative_TPC = TPC / sum(TPC)) %>% 
  ggplot(aes(x=Month, y=relative_TPC, group=State)) +
  geom_line(alpha = 0.08) +
  stat_summary(mapping = aes(x=Month,  y=relative_TPC, group = 1), 
               fun = "mean", 
               geom = "line",
               inherit.aes = FALSE) +
  ylab("Trips (as a fraction of annual total)") +
  ggtitle("Private Boats")
```

Figure \@ref(fig:private-boat-trips-seasonality-plot2) shows the same curves as 
in Figure \@ref(fig:private-boat-trips-seasonality-plot), but rescaled to show
values as a fraction of the annual total. This shows that fishing activity
is in most states fairly broadly distributed from April through November, with 
only 10 state-months accounting for more than 20% of the 
annual trips in a state, and a maximum of 35% for Maine in June. All but 1 of the 10
state-months with more than 20% of annual trips come from the Northern states of
Maine, New Hampshire, Massachusetts, Delaware and Rhode Island, with July in 
Mississippi as the exception.

```{r pTrips,echo=FALSE,warning=FALSE,fig.cap="2020 Monthly Marine Sportfishing Trips per Capita by State",fig.height=6,fig.width=5,fig.align='center'}
tripsTMPS %>%
  filter(Mode == 5 & Year > 2019) %>%
  ggplot( aes(x=Month,y=TPC, group=State, color=State)) +
  geom_line() + 
  ylab("Trips per capita") +
  ggtitle("Charter Boats")

tripsTMPS %>%
  filter(Mode == 3 & Year > 2019) %>%
  ggplot( aes(x=Month,y=TPC, group=State, color=State)) +
  geom_line() + 
  ylab("Trips per capita") +
  ggtitle("Shore")
```

```{r pString,echo=FALSE,warning=FALSE,fig.cap="Monthly 2020 stringency Index by State",fig.height=6,fig.width=5,fig.align='center'}
tripsTMPS %>% 
  filter(Year > 2019) %>%
  ggplot( aes(x=Month,y=Stringency, group=State, color=State)) +
  geom_line() + 
  ylab("stringency")

```

```{r tripsVstring,echo=FALSE,warning=FALSE,fig.cap="Average Trips per Capita versus Covid-19 Policy Stringency",fig.height=6,fig.width=5,fig.align='center'}

with(subset(mbms,Mode==7),
     plot(TPCdTPClt20 ~Stringency,pch=NA,main="Private Boats",xlim=c(1000,2000),
          ylab="Trips per Capita Ratio",xlab="Covid-19 Policy Stringency"))
with(subset(mbms,Mode==7 ),
     text(TPCdTPClt20 ~Stringency,font=2,cex=0.75,
          labels=State[Mode==7 ]))

with(subset(mbms,Mode==5 ),
     plot(TPCdTPClt20 ~Stringency,pch=NA,main="Charter Boats",xlim=c(1000,2000),
          ylab="Trips per Capita Ratio",xlab="Covid-19 Policy Stringency"))
with(subset(mbms,Mode==5 ),
     text(TPCdTPClt20 ~Stringency,font=2,cex=0.75,
          labels=State[Mode==5 ]))

with(subset(mbms,Mode==3 ),
     plot(TPCdTPClt20 ~Stringency,pch=NA,main="Shore",xlim=c(1000,2000),
          ylab="Trips per Capita Ratio",xlab="Covid-19 Policy Stringency"))
with(subset(mbms,Mode==3 ),
     text(TPCdTPClt20 ~Stringency,font=2,cex=0.75,
          labels=State[Mode==3 ]))

```



## Modeling Approach

Quasi-poisson regression estimated with mean equal to $exp(pop)*exp(xb)$ and variance equal to $scale*mean$.

# Results
Our results represent a composite effect of the actual local COVID policies, as well as localized attitudes, risk tolerances, and beliefs about COVID-19 which are correlated with COVID stringency. In particular, we see the inverse-U relationship we find to be the outcome of 2 countervailing pathways (laid out in the background section), (1) stringency reduces available substitutes of indoor or high-density outdoor activities, increasing the attractiveness of fishing, which is the dominant force on the low-stringency side of the curve, and (2) COVID risk makes even a relatively distanced activity, such as fishing, less appealing compared to entirely in-home activities such as cooking, backyard games, video games, etc., which is the dominating force on the high-stringency side of the curve.

```{r sumSat,echo=FALSE,warning=FALSE,results='asis',message=FALSE}

# ssbym=as.data.frame(t(aggregate(cbind(Trips,Hrsf,Party,Stringency)~Mode,data=tripsTMPS,
#           FUN=function(x) cbind(mean(x), median(x), sd(x)))))
# 
# names(ssbym)=c("Shore","Charter","Private")
# rownames(ssbym)=c("Mode",rep(c("Mean","Median","St. Dev."),4))
# 
# kbl(ssbym[-1,],row.names = rep(c("Mean","Median","St. Dev."),4)) %>%
#   kable_classic() %>%
#   pack_rows("Trips", 1, 3) %>%
#   pack_rows("Hours Fished", 4, 6)  %>%
#   pack_rows("Party Size", 7, 9)  %>%
#   pack_rows("Stringency", 10, 12)

library(vtable)
st(tripsTMPS[,c("Trips","Hrsf","Party","Stringency","Mode")],
   group = 'Mode', group.test = FALSE,group.long = TRUE,
   title="Summary Statistics by Mode",
   out="latex")

```



```{r pfe,echo=FALSE,warning=FALSE}

#Fixed effect poisson regressions of trips on stringency index with a population offset
#Modeling trips per capita
mf=formula(Trips ~ State + Month + I(Stringency/100) + I((Stringency/100)^2) + offset(log(Population)))
pm=glm(mf,data=tripsTMPS,family = quasipoisson,subset=Mode==7 )
cm=glm(mf,data=tripsTMPS,family = quasipoisson,subset=Mode==5 )
sm=glm(mf,data=tripsTMPS,family = quasipoisson,subset=Mode==3 )

```

```{r regTbl,echo=FALSE,warning=FALSE,results='asis'}

staten=gsub("State","",names(coef(pm)))[2:16]

texreg(list(pm,cm,sm),custom.model.names = c("Private","Charter","Shore"),single.row = FALSE,digits=3,label="regResults",
       caption = "Quasi-Poisson Fixed Effect Regression of Trips on Covid-19 Stringency by Mode",
       #custom.coef.names = c("Intercept",staten,month.name[-1],
       #                       "Stringency/100","(Stringency/100)^2"),
       custom.coef.map=list("(Intercept)" = "Intercept",
                            "I(Stringency/100)"="Stringency/100","I((Stringency/100)^2)"="(Stringency/100)^2"),
       include.aic=FALSE,include.bic=FALSE,include.loglik=FALSE,caption.above = TRUE,
       custom.note = "%stars. The base state is Alabama and the base Month is January.",
       custom.gof.rows = list("Scale" = c(summary(pm)$dispersion,summary(cm)$dispersion,summary(sm)$dispersion)))


```

```{r statesFEplot,echo=FALSE,warning=FALSE,fig.cap="State Fixed Effects for Trips by Mode",fig.height=8,fig.width=4,message=FALSE}

bn=list( "StateConnecticut" = "Connecticut",
         "StateDelaware" = "Delaware",
         "StateFlorida" = "Florida",
         "StateGeorgia" = "Georgia",
         "StateMaine" = "Maine", 
         "StateMaryland" = "Maryland",
         "StateMassachusetts" = "Massachusetts",
         "StateMississippi" = "Mississippi",
         "StateNew Hampshire" = "New Hampshire",
         "StateNew Jersey" = "New Jersey",
         "StateNew York" = "New York",
         "StateNorth Carolina" = "North Carolina",
         "StateRhode Island" = "Rhode Island",
         "StateSouth Carolina" = "South Carolina",
         "StateVirginia" = "Virginia",      
         "Month02" = "February",
         "Month03" = "March",
         "Month04" = "April",
         "Month05" = "May", 
         "Month06" = "June",
         "Month07" = "July",
         "Month08" = "August",
         "Month09" = "September",
         "Month10" = "October", 
         "Month11" = "November",
         "Month12" = "December" )

plotreg(list(pm,cm,sm),custom.model.names = c("Private","Charter","Shore"),
        custom.coef.map=bn[2:15])


```


```{r monthsFEplot,echo=FALSE,warning=FALSE,fig.cap="Months Fixed Effects for Trips by Mode",fig.height=8,fig.width=4,message=FALSE}

plotreg(list(pm,cm,sm),custom.model.names = c("Private","Charter","Shore"),
        custom.coef.map=bn[16:27])

```


```{r tripStringRel,echo=FALSE,warning=FALSE,fig.cap="Predicted Trips as a Function of the Stringency Index",fig.height=8,fig.width=4}

#Dataset to predict trips as a function of stringency
#holding everything else constant.
#Using June in Florida: zero stringency should correspond with
#the mean historical trips in Florida during June
nd=data.frame(State="Florida",Month="06",
              Population=with(tripsTMPS,mean(Population[State=="Florida" & Month=="06"])),
              Stringency=seq(0,3000,by=50))

#Predicted trips as a function of stringency, including confidence intervals
pd=add_ci(nd,pm)
cd=add_ci(nd,cm)
sd=add_ci(nd,sm)

#Plot predicted trips and confidence intervals
par(mfrow=c(3,1))

plot(y=pd$pred/1000000,x=pd$Stringency,type="l",ylim=c(2,5),
     main="Private Boats",
     xlab="Stringincy Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Millions of Fishing Trips (Scaled to June in FL)")
lines(y=pd$LCB0.025/1000000,x=pd$Stringency,lty=2)
lines(y=pd$UCB0.975/1000000,x=pd$Stringency,lty=2)

plot(y=cd$pred/1000000,x=cd$Stringency,type="l",ylim=c(.04,.09),
     main="Charter Boats",
     xlab="Stringincy Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Millions of Fishing Trips (Scaled to June in FL)")
lines(y=cd$LCB0.025/1000000,x=cd$Stringency,lty=2)
lines(y=cd$UCB0.975/1000000,x=cd$Stringency,lty=2)

plot(y=sd$pred/1000000,x=sd$Stringency,type="l",ylim=c(3,7),
     main="Shore",
     xlab="Stringincy Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Millions of Fishing Trips (Scaled to June in FL)")
lines(y=sd$LCB0.025/1000000,x=sd$Stringency,lty=2)
lines(y=sd$UCB0.975/1000000,x=sd$Stringency,lty=2)

par(mfrow=c(1,1))

#Use this to reverse the x-axis: xlim=c(3000,0)

#Don't think that we need robust standard error with the
#Quasi-poisson, but here is how you would calculate them
#coeftest(pm, vcov = vcovHC(pm, type="HC4"))

#Bootstrapped prediction CI to verify (takes long to run as currently coded)
# library(boot)
# 
# bf=function(d,i){
#   pm=glm(mf,data=d[i,],family = quasipoisson )
#   pd=predict(pm,newdata = nd,type="response")/1000000
#   return(pd)
# }
# 
# bcp=boot(subset(tripsTMPS,Mode==7),bf,R=1000)
# zz=sapply(1:61,FUN=function(i) boot.ci(boot.out = bcp, type = c("bca"),index=i)$bca)
# 
# plot(pd,type="l",ylim=c(2,6))
# lines(zz[4,],lty=2)
# lines(zz[5,],lty=2)

```

```{r lmfeh,echo=FALSE,warning=FALSE}

#Fixed effect regressions of mean hours fished on stringency index 
mfh=formula(Hrsf ~ State + Month + I(Stringency/100) + I((Stringency/100)^2))
pmh=lm(mfh,data=tripsTMPS,subset=Mode==7 )
cmh=lm(mfh,data=tripsTMPS,subset=Mode==5 )
smh=lm(mfh,data=tripsTMPS,subset=Mode==3 )

pmhr <- coeftest(pmh, vcov = vcovHC)
pse <- pmhr[, 2]
ppval <- pmhr[, 4]

cmhr <- coeftest(cmh, vcov = vcovHC)
cse <- cmhr[, 2]
cpval <- cmhr[, 4]

smhr <- coeftest(smh, vcov = vcovHC)
sse <- smhr[, 2]
spval <- smhr[, 4]

```

```{r regTblh,echo=FALSE,warning=FALSE,results='asis'}

texreg(list(pmh,cmh,smh),custom.model.names = c("Private","Charter","Shore"),single.row = FALSE,digits=3,label="regResultsh",
       caption = "Linear Fixed Effect Regression of Mean Monthly Hours Fished on Covid-19 Stringency by Mode",
       # custom.coef.names = c("Intercept",staten,month.name[-1],
       #                       "Stringency/100","(Stringency/100)^2"),
       custom.coef.map=list("(Intercept)" = "Intercept",
                            "I(Stringency/100)"="Stringency/100","I((Stringency/100)^2)"="(Stringency/100)^2"),
       include.aic=FALSE,include.bic=FALSE,include.loglik=FALSE,caption.above = TRUE,
       custom.note = ("\\parbox{.6\\linewidth}{\\vspace{2pt} %stars based on robust standard errors. \\\\
                      The base state is Alabama and the base Month is January.}"),
       override.se=list(pse,cse,sse),override.pvalues=list(ppval,cpval,spval))
```


```{r statesFEploth, echo=FALSE, fig.cap="State Fixed Effects for Hours Fished by Mode", fig.height=8, fig.width=4, warning=FALSE,message=FALSE}

plotreg(list(pmh,cmh,smh),custom.model.names = c("Private","Charter","Shore"),
        custom.coef.map=bn[2:15])


```


```{r monthsFEploth,echo=FALSE,warning=FALSE,fig.cap="Months Fixed Effects for Hours Fished by Mode",fig.height=8,fig.width=4,message=FALSE}

plotreg(list(pmh,cmh,smh),custom.model.names = c("Private","Charter","Shore"),
        custom.coef.map=bn[16:27])

```

```{r tripStringRelh,echo=FALSE,warning=FALSE,fig.cap="Predicted Hours Fished as a Function of the Stringency Index",fig.height=8,fig.width=4}

#Predicted trips as a function of stringency, including confidence intervals
pdh=add_ci(nd,pmh)
cdh=add_ci(nd,cmh)
sdh=add_ci(nd,smh)

#Plot predicted trips and confidence intervals
par(mfrow=c(3,1))

plot(y=pdh$pred,x=pdh$Stringency,type="l",ylim=c(3.5,5.5),
     main="Private Boats",
     xlab="Stringincy Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Mean Hours Fished (Scaled to June in FL)")
lines(y=pdh$LCB0.025,x=pdh$Stringency,lty=2)
lines(y=pdh$UCB0.975,x=pdh$Stringency,lty=2)

plot(y=cdh$pred,x=cdh$Stringency,type="l",ylim=c(3,7),
     main="Charter Boats",
     xlab="Stringincy Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Mean Hours Fished (Scaled to June in FL)")
lines(y=cdh$LCB0.025,x=cdh$Stringency,lty=2)
lines(y=cdh$UCB0.975,x=cdh$Stringency,lty=2)

plot(y=sdh$pred,x=sdh$Stringency,type="l",ylim=c(3,5.5),
     main="Shore",
     xlab="Stringincy Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Mean Hours Fished (Scaled to June in FL)")
lines(y=sdh$LCB0.025,x=sdh$Stringency,lty=2)
lines(y=sdh$UCB0.975,x=sdh$Stringency,lty=2)

par(mfrow=c(1,1))


```



```{r lmfep,echo=FALSE,warning=FALSE}

#Fixed effect regressions of mean hours fished on stringency index 
mfp=formula(Party ~ State + Month + I(Stringency/100) + I((Stringency/100)^2))
pmp=lm(mfp,data=tripsTMPS,subset=Mode==7 )
cmp=lm(mfp,data=tripsTMPS,subset=Mode==5 )

pmpr <- coeftest(pmp, vcov = vcovHC)
psep <- pmpr[, 2]
ppvalp <- pmpr[, 4]

cmpr <- coeftest(cmp, vcov = vcovHC)
csep <- cmpr[, 2]
cpvalp <- cmpr[, 4]

```

```{r regTblp,echo=FALSE,warning=FALSE,results='asis'}

texreg(list(pmp,cmp),custom.model.names = c("Private","Charter"),single.row = FALSE,digits=3,label="regResultsp",
       caption = "Linear Fixed Effect Regression of Mean Fishing Party Size on Covid-19 Stringency by Mode",
       # custom.coef.names = c("Intercept",staten,month.name[-1],
       #                       "Stringency/100","(Stringency/100)^2"),
       custom.coef.map=list("(Intercept)" = "Intercept",
                            "I(Stringency/100)"="Stringency/100","I((Stringency/100)^2)"="(Stringency/100)^2"),
       include.aic=FALSE,include.bic=FALSE,include.loglik=FALSE,caption.above = TRUE,
       custom.note = ("\\parbox{.6\\linewidth}{\\vspace{2pt} %stars based on robust standard errors. \\\\
                      The base state is Alabama and the base Month is January.}"),
       override.se=list(psep,csep),override.pvalues=list(ppvalp,cpvalp))
```


```{r statesFEplotp,echo=FALSE,warning=FALSE,fig.cap="State Fixed Effects for Party Size by Mode",fig.height=8,fig.width=4,message=FALSE}

plotreg(list(pmp,cmp),custom.model.names = c("Private","Charter"),
        custom.coef.map=bn[2:15])


```


```{r monthsFEplotp,echo=FALSE,warning=FALSE,fig.cap="Months Fixed Effects for Party Size Fished by Mode",fig.height=8,fig.width=4,message=FALSE}

plotreg(list(pmp,cmp),custom.model.names = c("Private","Charter"),
        custom.coef.map=bn[16:27])

```

```{r tripStringRelp,echo=FALSE,warning=FALSE,fig.cap="Predicted Party Size as a Function of the Stringency Index",fig.height=8,fig.width=4}

#Predicted party size as a function of stringency, including confidence intervals
pdp=add_ci(nd,pmp)
cdp=add_ci(nd,cmp)

#Plot predicted party size and confidence intervals
par(mfrow=c(2,1))

plot(y=pdp$pred,x=pdp$Stringency,type="l",ylim=c(2,5),
     main="Private Boats",
     xlab="Stringincy Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Mean Party Size (Scaled to June in FL)")
lines(y=pdp$LCB0.025,x=pdp$Stringency,lty=2)
lines(y=pdp$UCB0.975,x=pdp$Stringency,lty=2)

plot(y=cdp$pred,x=cdp$Stringency,type="l",ylim=c(3,7),
     main="Charter Boats",
     xlab="Stringincy Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Mean Party Size ( (Scaled to June in FL)")
lines(y=cdp$LCB0.025,x=cdp$Stringency,lty=2)
lines(y=cdp$UCB0.975,x=cdp$Stringency,lty=2)


par(mfrow=c(1,1))


```





```{r lmfem,echo=FALSE,warning=FALSE}

#Fixed effect regressions of mean hours fished on stringency index 
mfm=formula(Miles ~ State + Month + I(Stringency/100) + I((Stringency/100)^2))
pmm=lm(mfm,data=tripsTMPS,subset=Mode==7 )
cmm=lm(mfm,data=tripsTMPS,subset=Mode==5 )
smm=lm(mfm,data=tripsTMPS,subset=Mode==3 )

pmmr <- coeftest(pmm, vcov = vcovHC)
psem <- pmmr[, 2]
ppvalm <- pmmr[, 4]

cmmr <- coeftest(cmm, vcov = vcovHC)
csem <- cmmr[, 2]
cpvalm <- cmmr[, 4]

smmr <- coeftest(smm, vcov = vcovHC)
ssem <- smmr[, 2]
spvalm <- smmr[, 4]

```

```{r regTblm,echo=FALSE,warning=FALSE,results='asis'}

texreg(list(pmm,cmm,smm),custom.model.names = c("Private","Charter","Shore"),single.row = FALSE,digits=3,label="regResultsh",
       caption = "Linear Fixed Effect Regression of Mean Miles Traveled on Covid-19 Stringency by Mode",
       # custom.coef.names = c("Intercept",staten,month.name[-1],
       #                       "Stringency/100","(Stringency/100)^2"),
       custom.coef.map=list("(Intercept)" = "Intercept",
                            "I(Stringency/100)"="Stringency/100","I((Stringency/100)^2)"="(Stringency/100)^2"),
       include.aic=FALSE,include.bic=FALSE,include.loglik=FALSE,caption.above = TRUE,
       custom.note = ("\\parbox{.6\\linewidth}{\\vspace{2pt} %stars based on robust standard errors. \\\\
                      The base state is Alabama and the base Month is January.}"),
       override.se=list(psem,csem,ssem),override.pvalues=list(ppvalm,cpvalm,spvalm))
```


```{r statesFEplotm, echo=FALSE, fig.cap="State Fixed Effects for Miles Traveled by Mode", fig.height=8, fig.width=4, warning=FALSE,message=FALSE}

plotreg(list(pmm,cmm,smm),custom.model.names = c("Private","Charter","Shore"),
        custom.coef.map=bn[2:15])


```


```{r monthsFEplotm,echo=FALSE,warning=FALSE,fig.cap="Months Fixed Effects for Distance Traveled by Mode",fig.height=8,fig.width=4,message=FALSE}

plotreg(list(pmm,cmm,smm),custom.model.names = c("Private","Charter","Shore"),
        custom.coef.map=bn[16:27])

```

```{r tripStringRelm,echo=FALSE,warning=FALSE,fig.cap="Predicted Distance Traveled as a Function of the Stringency Index",fig.height=8,fig.width=4}

#Predicted trips as a function of stringency, including confidence intervals
pdm=add_ci(nd,pmm)
cdm=add_ci(nd,cmm)
sdm=add_ci(nd,smm)

#Plot predicted trips and confidence intervals
par(mfrow=c(3,1))

plot(y=pdm$pred,x=pdm$Stringency,type="l",ylim=c(20,35),
     main="Private Boats",
     xlab="Stringincy Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Mean Distance Traveled (Scaled to June in FL)")
lines(y=pdm$LCB0.025,x=pdm$Stringency,lty=2)
lines(y=pdm$UCB0.975,x=pdm$Stringency,lty=2)

plot(y=cdm$pred,x=cdm$Stringency,type="l",ylim=c(20,70),
     main="Charter Boats",
     xlab="Stringincy Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Mean Distance Traveled (Scaled to June in FL)")
lines(y=cdm$LCB0.025,x=cdm$Stringency,lty=2)
lines(y=cdm$UCB0.975,x=cdm$Stringency,lty=2)

plot(y=sdm$pred,x=sdm$Stringency,type="l",ylim=c(20,70),
     main="Shore",
     xlab="Stringincy Index (Monthly sum of daily records: higher is more stringent)",
     ylab="Mean Distance Traveled (Scaled to June in FL)")
lines(y=sdm$LCB0.025,x=sdm$Stringency,lty=2)
lines(y=sdm$UCB0.975,x=sdm$Stringency,lty=2)

par(mfrow=c(1,1))


```




# References {#references .unnumbered}
