---
title: "Importing updated COVID Stringency data"
output: html_notebook
---

```{r}
library(tidyverse)
```


We would like to consider updating our analysis to include 2021 data, and to
that end we need updated COVID stringency data, as the current file only goes
to April 2021. Unfortunately, the `xlsx` format data we have has not been updated, and there is a
csv we need to import:

```{r}
stringency_latest <- read_csv("OxCGRT_US_latest.csv")
```

```{r}
names(stringency_latest)
```

Let me just grab the columns we care about:

```{r}
stringency_latest <- stringency_latest %>% 
  select(CountryName, CountryCode, RegionName, RegionCode, Jurisdiction, Date,
         StringencyIndex, StringencyLegacyIndex) 

stringency_latest %>% head()
```

Let me re-format some of these columns:

But first check what values there are for jurisdiction...

```{r}
stringency_latest %>% 
  select(Jurisdiction) %>% distinct()
```

And values for RegionName:

```{r}
stringency_latest %>% 
  select(RegionName) %>% distinct()
```

It has the 50 states, plus DC and some NA values.

```{r}
stringency_latest <- stringency_latest %>% 
  rename(State = RegionName) %>% 
  mutate(Year = as.integer((substr(Date, start = 1, stop = 4))),
         Month = substr(Date, start = 5, stop = 6),
         Day = substr(Date, start = 7, stop = 8)) %>% 
  select(State, Year, Month, Day, StringencyIndex, StringencyLegacyIndex)

stringency_latest %>% head()
```

Lets collapse to the month level and compare values:

```{r}
stringency_monthly <- stringency_latest %>% 
  group_by(State, Year, Month) %>% 
  summarise(Stringency = sum(StringencyIndex),
            StringencyLegacy = sum(StringencyLegacyIndex),
            .groups = "drop")

stringency_monthly
```

These values are, unfortunately, not matching up with the values from the prior
data. However, I have figured out the problem. The old values we had were the sum
of the state-specific indices and the national value.

I need to get the national values of stringency for each date:

```{r}
national_stringency <- stringency_latest %>% 
  filter(is.na(State)) %>% 
  select(Year, Month, Day, StringencyIndex) %>% 
  rename(NationalStringencyIndex = StringencyIndex)

national_stringency
```

I can now add these to the state values for each state, and collapse to the monthly
level:

```{r}
stringency_monthly <- stringency_latest %>% 
  filter(!is.na(State)) %>% 
  left_join(national_stringency, by = c("Year", "Month", "Day")) %>% 
  mutate(StringencyIndex = StringencyIndex + NationalStringencyIndex) %>% 
  group_by(State, Year, Month) %>% 
  summarise(Stringency = sum(StringencyIndex),
            .groups = "drop")

stringency_monthly
```


This also does not match up with the old data. After some extensive investigation,
I have figured out the problem. There are 2 versions of the COVID stringnecy
index for US states, and I am looking at the wrong one, or at least a different
one than we had for the old data.

The data I have downloaded and investigated is the one here:

https://github.com/OxCGRT/USA-covid-policy

This is a USA-specific data repo

And the data that matches the old data we have been using for the paper is here:

https://github.com/OxCGRT/covid-policy-tracker


The difference between the stringency indices provided by the files in these 2 
different repos (general repo and US repo) is that the general repo accounts for
national policies in calculating the state-specific stringency index, while the US
specific repo does not. In particular, the general repo replaces component scores
for state-days where the national policy on that day is more stringent than the
state policy. For instance, for the travel indicator component of the index,
if Alabama had a 2 and Maine had 4, and the US policy was a 3, then under the general
repos methodology Alabama would have a 3 and Maine would have a 4, while under the US
repo's methodology, Alabama would retain its 2. 

In order to be consistent with the old data, we need to use the general repo's
data. However, I wonder if this strategy is not masking some meaningful variation
in state policies, especially considering that US Federal policy was certainly not 
meaningfully adhered to in many states.... It may be worth using both versions of 
the index and comparing results, and also I should dig more into the discrepancies
and try to reason through which methodology is more appropriate based on that.

```{r}
stringency_global <- read_csv("OxCGRT_latest.csv")
```

Get just the US state-specific data:


```{r}
stringency_us <- stringency_global %>% 
  filter(CountryCode == "USA") %>% 
  rename(State = RegionName) %>% 
  mutate(Year = as.integer((substr(Date, start = 1, stop = 4))),
         Month = substr(Date, start = 5, stop = 6),
         Day = substr(Date, start = 7, stop = 8)) %>% 
  select(State, Year, Month, Day, StringencyIndex)

stringency_us
```

Collapse to monthly level:

```{r}
stringency_us_monthly <- stringency_us %>% 
  group_by(State, Year, Month) %>% 
  summarise(Stringency = sum(StringencyIndex),
            .groups = "drop")

stringency_us_monthly
```

Let me compare these with the old values:

```{r}
tripsTMPS_p <- readRDS("tripsTMPS.rds") %>% filter(Mode == "7")

stringency_us_monthly %>% 
  left_join(tripsTMPS_p, by = c("State", "Year", "Month")) %>% 
  filter(!is.na(Stringency.y)) %>% 
  filter(Stringency.x != Stringency.y) %>% 
  select(State, Year, Month, Stringency.x, Stringency.y)
```

There are 40 state-months where there is some discrepancy between this updated data, and the
old data we had been using, mostly in February and March. Specifically:

```{r}
stringency_us_monthly %>% 
  left_join(tripsTMPS_p, by = c("State", "Year", "Month")) %>% 
  filter(!is.na(Stringency.y)) %>% 
  filter(Stringency.x != Stringency.y) %>% 
  select(State, Year, Month, Stringency.x, Stringency.y) %>% 
  count(Month)
```

80% of these discrepancies are from Feb and March 2020, with others randomly throughout
the year.

Plotting:

```{r}
stringency_us_monthly %>% 
  left_join(tripsTMPS_p, by = c("State", "Year", "Month")) %>% 
  filter(!is.na(Stringency.y)) %>% 
  filter(Stringency.x != Stringency.y) %>% 
  select(State, Year, Month, Stringency.x, Stringency.y) %>% 
  ggplot(aes(Stringency.x, Stringency.y)) + geom_point()
```

These are clearly very minor discrepancies, and I feel fine treating this new
data as essentially the same as the old data for all practical purposes.

Let me sum up with all the code needed to get to the monthly dataset:

```{r}
read_csv("OxCGRT_latest.csv") %>%
  filter(CountryCode == "USA") %>% 
  rename(State = RegionName) %>% 
  mutate(Year = as.integer((substr(Date, start = 1, stop = 4))),
         Month = substr(Date, start = 5, stop = 6)) %>% 
  group_by(State, Year, Month) %>% 
  summarise(Stringency = sum(StringencyIndex),
            .groups = "drop")
```






















