---
title: "Calculating and Checking SE's"
output: html_notebook
---

I can load the file that David created to calculate standard errors, in addition
to loading the normal file that we have been using to do the regression analysis.

```{r}
tripsTMPS <- readRDS("../tripsTMPS.rds")
tripsTs <- readRDS("../tripsTs.rds")
```

```{r}
library(tidyverse)
```


One snag already is that it seems the file David sent is the one in which he 
has already removed state-months that were not sampled. Let me verify:

```{r}
tripsTs %>% 
  count(ST, month) %>% 
  pivot_wider(names_from = month, values_from = n)
```

Yes, it seems that is exactly what has happened: we are missing observations from
all state-months that were not sampled, according to the chart in the MRIP
documentation which David used to define the not sampled variable. So it should
be true that this variable only has 1 value in the current data:

```{r}
tripsTs %>% 
  count(notSampled)
```

Indeed.

Let me double check that the values from these datasets at least line up
properly.

But first let me ensure there are no NA's in the trips from the old dataset:

```{r}
tripsTMPS %>% 
  filter(is.na(Trips))
```

Great. Now, what is the largest difference among values which do appear in both
datasets?

```{r}
tripsTs_forjoin <- tripsTs %>% 
  mutate(YEAR = as.character(YEAR) %>% as.numeric(),
         MODE_FX = as.character(MODE_FX),
         ST = as.character(ST) %>% as.numeric(),
         month = as.character(month))


trips_joined <- tripsTMPS %>% 
  left_join(tripsTs_forjoin, by = c("FIPS" = "ST", "Year" = "YEAR", 
                                    "Mode" = "MODE_FX", "Month" = "month"))

trips_joined %>% 
  filter(Trips != one) %>% 
  select(Year, Month, Mode, State, Trips, one) %>% 
  mutate(trip_diff = Trips - one) %>% 
  arrange(trip_diff)
```

The differences are all minute. But of course there are some values of NA in the
new data. Let me check if the old data has any non-0 values there:

```{r}
trips_joined %>% 
  filter(is.na(one), Trips > 0)
```

Nope! So, the dependent variable has the same information between these 2 datasets.
Great. I think I can basically work with the joined dataset, dropping the "one"
variable and focusing on the PSE's.

I would also like to know if there are any state months that I am dropping
when I use the old code for dropping 0 state months, which we are not dropping from
this current dataset.

```{r}
no_data_sms<- tripsTMPS %>% 
  filter(Mode == "7") %>% 
  group_by(State, Month) %>% 
  summarise(max_Trips = max(Trips), .groups = "drop") %>% 
  filter(max_Trips == 0) %>% 
  select(State, Month)

no_data_sms %>% 
  mutate(one=1) %>% 
  pivot_wider(names_from = Month, values_from = one)
```

So the difference I can see here is that my code also removed March for Mass and
RI. 

And then what about for the charter mode:

```{r}
no_data_sms5<- tripsTMPS %>%
  filter(Mode == "5") %>%
  group_by(State, Month) %>%
  summarise(max_Trips = max(Trips), .groups = "drop") %>%
  filter(max_Trips == 0) %>%
  select(State, Month)

no_data_sms5 %>% 
  mutate(one=1) %>% 
  pivot_wider(names_from = Month, values_from = one) %>% 
  relocate(`03`, .after = `02`)
```

This seems to be telling me for instance that Mass has 0 trips for charter mode in
all years for November. Let me look at the data to see if that lines up:

```{r}
trips_joined %>% 
  filter(Mode == "5", State == "Massachusetts", Month == "11")
```


Ok, so at least things line up, that there is no estimated SE here either way. 
Is there any way to programatically confirm that? And if so, am I correct in thinking
that this means that both datasets are essentially handling this missing data problem
in the same way? That is, since there are no observations in the intercept data for
Mass in November for charter mode, it means that we can not estimate the mean
number of trips, and so it needs to be excluded from estimation whether or not
the MRIP design intentionally doesn't sample that state month?

Ok, now what about Connecticut in March and April?

```{r}
trips_joined %>% 
  filter(Mode == "5", State == "Connecticut", Month %in% c("03", "04"))
```

There is only a single year in which Conn has data for March, which is 2018.
In all other years, there is no data for it. I can't help but wonder
if this means that I should be dropping those observations from the data.

How about this, how many more state-months do I drop if I demand that all years
have a minimum above 0, rather than just demanding that the maximum across all years
be greater than 0?

```{r}
some_no_data_sms<- tripsTMPS %>% 
  filter(Mode == "7") %>% 
  group_by(State, Month) %>% 
  summarise(min_Trips = min(Trips), .groups = "drop") %>% 
  filter(min_Trips == 0) %>% 
  select(State, Month)

some_no_data_sms %>% 
  mutate(one=1) %>% 
  pivot_wider(names_from = Month, values_from = one)
```

Ok, so this is only a slightly more expansive list, with an extra 4 state-months.

Let me look at the values obtained after filtering out all of these state-months:

```{r}
trips_joined_7_valid <- trips_joined %>% 
  anti_join(some_no_data_sms, by = c("State", "Month")) %>% 
  filter(Mode == "7")
```

```{r}
trips_joined_7_valid$Trips %>% min
```

Great, so for the private mode, once I filter down like this, there are no observations
with 0 trips. So it should also be the case that there are no NA's on the `one` variable:

```{r}
trips_joined_7_valid %>% 
  filter(is.na(one))
```

great. And what about the standard errors, are any of these missing?

```{r}
trips_joined_7_valid %>% 
  arrange(pse)
```

It seems that I do have just a few observations, all from delaware and Conn in 
March, for which the PSE is showing up as 0. This presumably means that there
is only 1 observation for them, and so SE's were not able to be calculated.


How does all of this look for the charter mode?

```{r}
some_no_data_sms5<- tripsTMPS %>% 
  filter(Mode == "5") %>% 
  group_by(State, Month) %>% 
  summarise(min_Trips = min(Trips), .groups = "drop") %>% 
  filter(min_Trips == 0) %>% 
  select(State, Month)

some_no_data_sms5 %>% 
  mutate(one=1) %>% 
  pivot_wider(names_from = Month, values_from = one)
```

So in fact this condition becomes much more restrictive for the charter mode.
Perhaps it would still be worth trying it out and seeing how it changes
the results of the paper, and whether it might still be worth keeping it this way
for the sake of having things a little cleaner.

Let me double check that things are fine with the shore mode as well:

```{r}
trips_joined_3_valid <- trips_joined %>% 
  anti_join(some_no_data_sms, by = c("State", "Month")) %>% 
  filter(Mode == "3")

trips_joined_3_valid$Trips %>% min

trips_joined_3_valid %>% 
  filter(is.na(one))

trips_joined_3_valid %>% 
  arrange(pse)
```


So actually for the shore mode I need to drop Conn in March if I am going to take this approach.

It is worth adding all of this code to the paper, in an appendix. Then I can re-define
the function for doing the regressions models to specifically handle the shore mode
as well. And I can make sure that each mode's regression model has separate datasets.














