---
title: "Compare new and old datasets"
output: html_notebook
---

I want to verify that the new dataset I have produced with 2021 data matches 
the previous dataset for the observations through 2020.

```{r}
library(tidyverse)
```


```{r}
tripsTMPS_new <- readRDS("tripsTMPS_new.rds")
tripsTMPS_old <- readRDS("../tripsTMPS.rds")

tripsTMPS_new
tripsTMPS_old
```


So, the old dataset has 2 variables (Miles and Htvl) which I did not bother
to include in the new data, since we decided not to use those outcomes. I also
should sort both the files on the same way:

```{r}
tripsTMPS_new <- tripsTMPS_new %>% 
  arrange(Year, Month, Mode, State)

tripsTMPS_old <- tripsTMPS_old %>% 
  select(!c(Miles,Htvl)) %>% 
  arrange(Year, Month, Mode, State) %>% 
  mutate(FIPS = as.numeric(FIPS)) %>% 
  as_tibble()

tripsTMPS_new
tripsTMPS_old
```


There are still some discrepancies, let me go back and fix the new file.


Ok, the files are now largely identical, let me identify the discrepancies:

```{r}
tripsTMPS_new_2020 <- tripsTMPS_new %>% filter(Year <= 2020)

dsicrepancies_new_values <- tripsTMPS_new_2020 %>% 
  anti_join(tripsTMPS_old)

dsicrepancies_new_values
```

There are 123 observations which are not identical (`anti_join` shows all rows in
x which are not in y, and here since all the columns match, it is showing rows
that are not identical), let me compare them with their old values:

```{r}
dsicrepancies_new_values %>% 
  left_join(tripsTMPS_old, by = c("Year", "Month", "Mode", "State")) %>% 
  mutate(Stringency_diff = abs(Stringency.x - Stringency.y)) %>% 
  arrange(Stringency_diff)
```

Here we can see that there are some discrepancies in the value of stringency used,
as I already knew would be the case. I am not too concerned about this, as the 
difffernces are small, and the large majority of observations have no discrepancies
at all.

























