---
title: "Figure out independent variable"
output: html_notebook
---

We are interested in determining what is an appropriate independent variable for
this analysis. We have so far been using the Stringency Index, but are now 
considering whether it would be more appropriate to use one of the components
rather than the entire index, since it is not a priori clear what should be the
impact of many of the components on trips. We have in particular honed in on
the "stay at home order" component of the index as a good candidate for our
independent variable.

But, in the first place, we need to determine what is the difference between
2 different versions of the dataset.

```{r}
library(tidyverse)
```

```{r}
stringency_us_latest <- read_csv("../data/OxCGRT_US_latest.csv") %>% 
  rename(State = RegionName) %>% 
  mutate(Year = as.integer((substr(Date, start = 1, stop = 4))),
         Month = substr(Date, start = 5, stop = 6),
         Day = substr(Date, start = 7, stop = 8)) %>% 
  arrange(State, Year, Month, Day)


stringency_latest <- read_csv("../data/OxCGRT_latest.csv") %>% 
  filter(CountryCode == "USA") %>% 
  rename(State = RegionName) %>% 
  mutate(Year = as.integer((substr(Date, start = 1, stop = 4))),
         Month = substr(Date, start = 5, stop = 6),
         Day = substr(Date, start = 7, stop = 8)) %>% 
  arrange(State, Year, Month, Day)
```


I want to look at the values of the `C6_Stay at home requirements` variable,
and see if there are any discrepancies between these data, and if so for what dates.

I can do this by joining the 2 datasts, and then looking at the 2 different versions
of the variables from each dataset:

```{r}
stringency_latest %>% 
  left_join(stringency_us_latest, by = c("State", "Year", "Month", "Day")) %>% 
  filter(`C6_Stay at home requirements.x` != `C6_Stay at home requirements.y`) %>% 
  select(State, Year, Month, Day, `C6_Stay at home requirements.x`, `C6_Stay at home requirements.y`) %>% 
  arrange(State, Year, Month, Day) %>% 
  nrow()

stringency_latest %>% 
  left_join(stringency_us_latest, by = c("State", "Year", "Month", "Day")) %>% 
  filter(`C6_Stay at home requirements.x` > `C6_Stay at home requirements.y`) %>% 
  count(State, Year, Month)

stringency_latest %>% 
  left_join(stringency_us_latest, by = c("State", "Year", "Month", "Day")) %>% 
  filter(`C6_Stay at home requirements.x` < `C6_Stay at home requirements.y`) %>% 
  count(State, Year, Month)
```

I am finding the the general file has many observations where the stay at home 
orders are considered more strict, when compared to the US-specific file. Whereas
there are only a few observations in 2022 for Alaska where the US-specific file
has more stringent policies listed. This seems to reverse my understanding of
how the data works, as I thought that the US-specific policy variables file
incorporated national level policies, while the other did not.

There seems to be something happening in April 2021, as this is where many of the
states cease to have any observations with discrepancies.

Looking at the notes for national policies, it seems there was a change in national
"policy" on Apr 2 2021: On April 2, recommendations on the CDC's Travel During COVID-19 site were updated from "Delay travel and stay home to protect yourself and others from COVID-19" to not explicitly recommending stay at home. Additionally, the site was updated with the recommendation that fully vaccinated individuals can safely travel. 

Therefore, the data from the general repo continues to have a `1` for most states
up until Apr 2, while in the US data those states cease to have `1` many months before
Apr 2021. This however would not make a difference to the variable we have considered
constructing of how many days the state is at a `2`.

























