---
title: "Create Origin-Destination dataset"
output: html_notebook
---

I want to create a version of the data in which the trips are summarized
not only by the state in which they occurred, but also by the state the anglers
are from.

```{r}
library(feather)
library(survey)
options(survey.lonely.psu="certainty",multicore=TRUE)
library(cdlTools)
library(tidyverse)
```

```{r}
#Read the trips data
trips <- read_feather("../data/trips1721.feather")
```

```{r}
# add state abbreviations
trips$STa=fips(trips$ST,to="Abbreviation")
trips$ST_RESa=fips(trips$ST_RES,to="Abbreviation")

with(subset(trips,MODE_FX==5),table(STa,WAVE))
```

Here I will use the survey package to come up with the estimates of trips for
each origin-destination pair of states. This will include standard errors for
the estimates.

```{r}
#Add counter varible
trips$one=1
#Trips survey design object
tripss=svydesign(id=~psu_id, strata=~strat_id,weights=~wp_int,data=trips,nest=TRUE)
#Estimate total trip counts by year,month,mode, and state, including SEs
tripsT=svyby(~one,~YEAR+month+MODE_FX+ST+ST_RES,FUN=svytotal,
             design=subset(tripss,wp_int > 0 & month != "99" & MODE_FX != "4"),
             drop.empty.groups=FALSE,multicore = TRUE)
```

I can also come up with the same estimates by simply taking the sum of the survey
weights, and that way I can also get the count of how many survey records were
included in each of those cells.


```{r}
tripsTM <- trips %>% 
  filter(wp_int > 0, month != 99, MODE_FX != 4) %>% 
  # exclude anomalous observations
  complete(YEAR,month,MODE_FX,ST, ST_RES) %>%
  group_by(YEAR,month,MODE_FX,ST, ST_RES) %>%
  summarize(N_obs = n(),
            Trips = sum(wp_int,na.rm=TRUE),
            .groups = "drop") %>% 
  mutate(MONTH = as.numeric(month),
         year = as.numeric(as.character(YEAR)),
         fips = as.numeric(as.character(ST)))
```

```{r}
tripsTM %>% nrow()
# get total unique combinations of origin-destination pairs
tripsTM %>% distinct(ST, ST_RES) %>% nrow()
# get number of combinations that have positive trips
tripsTM %>% filter(Trips > 0) %>% distinct(ST, ST_RES) %>% nrow()
# find combinations which have observations for every mode, month and year
tripsTM %>% filter(Trips>0) %>% count(ST, ST_RES) %>% filter(n==180)
```


With 3 modes, 12 months, 5 years, 16 destination states, and 63 origin states/territories,
there are 181,440 total possible observations in this dataset. However, of the
1,008 unique origin-destination pairs in the data, many do not have any observations
in the dataset for a given mode, month, and year. 216 of the pairs do not
have observations in any mode, month, or year. Only one of the pairs has observations
for every single one of the 180 mode-month-years, and that is Florida residents taking
trips in Florida. The remaining origin-destination pairs will all have some amount
of missing data. 

```{r}
tripsTM %>% filter(Trips>0) %>% nrow()
```

Of the 181,440 possible observations, only 21,824 have any survey data to inform
the estimate of number of trips.

I need to add the population data as well as stringency data for both origin
and destination states. The population for each observation should be
the population of the origin state.

```{r}
pop2019 <- tidycensus::get_acs(geography = "state", 
            variables = "B01001_001", 
            year = 2019,
            survey = "acs1") 
```



```{r}
pop2019 <- pop2019 %>% 
  mutate(fips = as.numeric(GEOID))

tripsTMP <- tripsTM %>% 
  left_join(pop2019, by = c("ST_RES"="fips")) %>% 
  rename(population = estimate) %>% 
  select(-c(GEOID, moe, NAME, variable))
```


The population for the origin state is now added. How many states actually have
population data?

```{r}
tripsTMP %>% distinct(ST_RES) %>% nrow()
tripsTMP %>% filter(!is.na(population)) %>% distinct(ST_RES) %>% nrow()
```

52, as it should be.

```{r}
fips_codes_state <- tidycensus::fips_codes %>% select(state_code, state_name) %>% distinct()

stringency <- read_csv("../data/OxCGRT_US_latest.csv") %>% 
  rename(State = RegionName) %>% 
  mutate(Year = as.integer((substr(Date, start = 1, stop = 4))),
         Month = substr(Date, start = 5, stop = 6),
         Day = substr(Date, start = 7, stop = 8)) %>% 
  # next 2 commands assign the last known value of Stringency for days with NA
  arrange(State, Year, Month, Day) %>% 
  fill(StringencyIndex) %>% 
  group_by(State, Year, Month) %>%
  summarise(Stringency = mean(StringencyIndex),
            .groups = "drop") %>% 
  left_join(fips_codes_state, by = c("State" = "state_name")) %>% 
  rename(fips = state_code) %>% 
  mutate(fips = as.integer(fips))
```


```{r}
tripsTMPS_od <- tripsTMP %>% 
  left_join(stringency, by = c("year" = "Year", "month" = "Month", "ST" = "fips")) %>% 
  rename(state_destination = State, stringency_destination = Stringency) %>% 
  left_join(stringency, by = c("year" = "Year", "month" = "Month", "ST_RES" = "fips")) %>% 
  rename(state_origin = State, stringency_origin = Stringency) %>% 
  select(-c(state_destination, state_origin)) %>% 
  filter(!is.na(population))
```

Now I have the data I need, and I have filtered down to only the 52 origin states
of the main US, I could ultimately probably trim even further but this is good
for now. 

Are there any missing values of stringency?

```{r}
tripsTMPS_od %>% filter(year>=2020, is.na(stringency_destination)) %>% nrow()
tripsTMPS_od %>% filter(year>=2020, is.na(stringency_origin)) %>% nrow()
```

There are missing values for both Puerto Rico and DC, I can remove these:

```{r}
tripsTMPS_od <- tripsTMPS_od %>% 
  filter(!ST_RES %in% c(11, 72))
```


```{r}
tripsTMPS_od %>% filter(year>=2020, is.na(stringency_origin)) %>% nrow()
```


And now I have no missing values for stringency in 2020 and later, and I can impute
stringency equal to zero for earlier years:

```{r}
tripsTMPS_od <- tripsTMPS_od %>% 
  mutate(stringency_destination = if_else(is.na(stringency_destination), 0, stringency_destination),
         stringency_origin = if_else(is.na(stringency_origin), 0, stringency_origin))
```

I should now be able to fit a model:

```{r}
library(fixest)
```

```{r}
trips_pois <- fepois(Trips ~ stringency_origin + stringency_destination | ST^ST_RES, tripsTMPS_od)
print(trips_pois)
```


Here I am using all of the observations, many of which are 0 only because of missing
data, let me restrict to those with positive trips and see how the results look.

```{r}

trips_pois2 <- fepois(Trips ~ stringency_origin + stringency_destination | ST^ST_RES, 
                      data = tripsTMPS_od %>% filter(Trips>0))
print(trips_pois2)
```


Here the signs of the coefficients are in the expected directions, but are not
estimated very precisely. I can add the offset to provide some more structure:

```{r}
trips_pois3 <- fepois(Trips ~ stringency_origin + stringency_destination | ST^ST_RES, 
                      data = tripsTMPS_od %>% filter(Trips>0),
                      offset = ~log(population))
print(trips_pois3)
```

This actually does not affect the results at all.

How many observations are there with more than 0 trips:

```{r}
tripsTMPS_od %>% filter(Trips>0) %>% nrow()
```

how many observations are there for each fixed effect?

```{r}
tripsTMPS_od %>% filter(Trips>0) %>% 
  count(ST, ST_RES) %>% 
  arrange(n)
```

of 730 unique combinations of origin and destination states, there are 87 which
have only a single observations, which should therefore be dropped. Let me see if
results are changed when specifying a different way of dropping observations:

```{r}
trips_pois4 <- fepois(Trips ~ stringency_origin + stringency_destination | ST^ST_RES, 
                      data = tripsTMPS_od %>% filter(Trips>0),
                      offset = ~log(population),
                      fixef.rm = "both")
print(trips_pois4)
```

This does affect the number of observations, but has no impact on the precision of
the estimates.

It will be good to save my dataset for future use:

```{r}
saveRDS(tripsTMPS_od, "../tripsTMPS_od.rds")
```

