---
title: "Changes in MRIP Survey Weights 2019 to 2020"
output: html_notebook
---

Here I will examine whether the pattern of MRIP survey weights changed substantially from 2019 
to 2020, to assuage concerns about the impact of COVID on the data collection.

```{r}
library(feather)
library(tidyverse)
```

```{r}
#Read the trips data
trips <- read_feather("../data/trips1721.feather")
```

The survey weight variable is `wp_int`. I want to create separate histograms
by year for this variable:

```{r}
trips %>% 
  filter(YEAR == 2019, wp_int < 20000) %>% 
  ggplot(aes(x=wp_int)) + 
  geom_histogram()
```

```{r}
trips %>% 
  filter(YEAR == 2020, wp_int < 20000) %>% 
  ggplot(aes(x=wp_int)) + 
  geom_histogram()
```


```{r}
trips %>% 
  filter(YEAR == 2018, wp_int < 20000) %>% 
  ggplot(aes(x=wp_int)) + 
  geom_histogram()
```


These look fairly similar to me, but perhaps I could also look at the log of the
survey weight variable to see the differences more clearly:

```{r}
trips %>% 
  mutate(log_survey_weight = log(wp_int)) %>% 
  filter(YEAR==2019) %>% 
  ggplot(aes(x=log_survey_weight)) +
  geom_histogram() + xlab("log survey weights")
```


```{r}
trips %>% 
  mutate(log_survey_weight = log(wp_int)) %>% 
  filter(YEAR==2020) %>% 
  ggplot(aes(x=log_survey_weight)) +
  geom_histogram() + xlab("log survey weights")
```













